
\textbf{Figure 1} (contamination_rates_barplot.png): 13-gram contamination rates across all 59 benchmark sub-tasks against The Pile v1, sorted in ascending order. Kruskal-Wallis H=590.82, p=2.73×10⁻⁸⁹. The distribution spans 0.4\% (high_school_mathematics) to 17.3\% (professional_medicine).

\textbf{Figure 2} (top_bottom.png): Top-10 most contaminated and bottom-10 least contaminated benchmark sub-tasks in The Pile v1. The 40× differential between professional_medicine (17.3\%) and high_school_mathematics (0.4\%) is visualized.

\textbf{Figure 3} (fig1_corpus_mean_comparison.png): Mean 13-gram contamination rates for the three training corpora. The Pile v1 (6.53\%) > RedPajama-v1 (5.75\%) > C4 en.noclean (4.05\%). Kruskal-Wallis H=17.51, p=1.58×10⁻⁴.

\textbf{Figure 4} (fig2_contamination_heatmap.png): Contamination heatmap for the top-20 most contaminated sub-tasks across 3 corpora. The corpus ordering (Pile ≥ RedPajama > C4) is visible across sub-tasks.

\textbf{Figure 5} (domain_corpus_heatmap.png): Domain-stratified contamination rates (2×3 heatmap: academic vs. commonsense × 3 corpora). Academic MMLU sub-tasks show consistently higher contamination than commonsense benchmarks across all three corpora. Cohen's d=0.85; exploratory result — see Section 5.3 for limitations.

\textbf{Figure 6} (sensitivity_scatter.png): Sensitivity analysis showing contamination rankings under question+choices vs. question-only text format. Spearman ρ=0.74, p=2.69×10⁻¹¹.

\textbf{Figure 7} (fig3_corpus_pair_scatter.png): Corpus-pair scatter plots showing per-sub-task contamination correlations across the three corpus pairs. The tight Pile-RedPajama cluster relative to the wider Pile-C4 spread is consistent with the Dunn post-hoc significance pattern.
