
\subsubsection{3.1 System Architecture}

The SVAD (Semantic Versioning with Adaptive Drift-Based Deprecation) system consists of three layers. This work evaluated the first two layers; the third was not implemented.

\textbf{Layer 1 - Detection:} For dataset version pair (v_old, v_new), the system extracted features using frozen pre-trained models (ResNet-50 for vision with 2048-dimensional embeddings, BERT-base for NLP with 768-dimensional embeddings). Features were reduced to 2 principal components using PCA. Drift was measured using:

\begin{itemize}
\item Kolmogorov-Smirnov test: maximum distance between empirical CDFs, averaged across dimensions
\item Maximum Mean Discrepancy: kernel-based distance using Gaussian kernel with bandwidth set via median heuristic
\end{itemize}

The final drift score was the maximum of the KS and MMD scores.

\textbf{Layer 2 - Classification:} Drift scores were compared against fixed thresholds:
\begin{itemize}
\item MAJOR: drift ≥ 0.07
\item MINOR: 0.02 ≤ drift < 0.07
\item PATCH: drift < 0.02
\end{itemize}

These thresholds were derived from Recht et al.'s ImageNet-v2 findings (11-14\% performance drop) scaled to approximately 7\% based on estimated PCA sensitivity loss.

\textbf{Layer 3 - Deprecation:} This layer was not implemented or evaluated.

\subsubsection{3.2 Feature Extraction}

Vision datasets used ResNet-50 \citep{Heetal2016} pre-trained on ImageNet, extracting 2048-dimensional features from the penultimate layer. NLP datasets used BERT-base \citep{Devlinetal2019} pre-trained on Wikipedia and BooksCorpus, extracting 768-dimensional [CLS] token embeddings. All model weights were frozen. PCA reduced features to 2 components.

\subsubsection{3.3 Statistical Tests}

The Kolmogorov-Smirnov statistic was computed as:

$$D_{KS} = \sup_x |F_{old}(x) - F_{new}(x)|$$

where $F_{old}$ and $F_{new}$ are empirical CDFs. The statistic was averaged across both PCA dimensions.

Maximum Mean Discrepancy was computed as:

$$\text{MMD}^2(P, Q) = \mathbb{E}_{x,x' \sim P}[k(x,x')] + \mathbb{E}_{y,y' \sim Q}[k(y,y')] - 2\mathbb{E}_{x \sim P, y \sim Q}[k(x,y)]$$

using Gaussian RBF kernel $k(x, y) = \exp(-\|x - y\|^2/(2\sigma^2))$ with bandwidth $\sigma$ set to the median pairwise distance.

\subsubsection{3.4 Evaluation Protocol}

Datasets were selected based on documented version histories. Ground truth labels (MAJOR/MINOR/PATCH) were assigned based on literature descriptions rather than measured model performance. The classifier's predictions were compared against these labels to compute accuracy, precision, recall, and F1 score.

Success criteria were predefined as:
\begin{itemize}
\item Overall accuracy ≥ 85\%
\item Precision for MAJOR changes ≥ 70\%
\item Recall for MAJOR changes ≥ 85\%
\end{itemize}

These criteria were established before conducting experiments.
