
\subsubsection{5.1 Primary Metrics}

Table 1 presents results against predefined success criteria:

\begin{table}[htbp]
\centering
\begin{tabular}{lllll}
\toprule
Metric & Target & Achieved & Gap & Status \\
\midrule
Accuracy & 85\% & 28.57\% & -56.43pp & Failed \\
Precision (MAJOR) & 70\% & 25\% & -45pp & Failed \\
Recall (MAJOR) & 85\% & 25\% & -60pp & Failed \\
F1 (MAJOR) & 75\% & 25\% & -50pp & Failed \\
\bottomrule
\end{tabular}
\end{table}

The classifier achieved 28.57\% overall accuracy, which is 4.73 percentage points below the 33\% random baseline for three classes. Precision and recall for MAJOR detection were both 25\%, substantially below targets.

\subsubsection{5.2 Confusion Matrix}

Table 2 shows the confusion matrix:

```
              Predicted
           MAJOR  MINOR  PATCH  Total
Actual:
MAJOR        1      2      1      4
MINOR        2      1      0      3
PATCH        4      2      1      7
```

The classifier correctly identified 1 out of 4 MAJOR changes (25\% recall), missing 3 MAJOR changes (2 classified as MINOR, 1 as PATCH). This represents a 75\% false negative rate for critical breaking changes.

Of 7 predictions labeled MAJOR, only 1 was correct (25\% precision). The remaining 6 were false positives: 4 PATCH datasets and 2 MINOR datasets incorrectly flagged as MAJOR.

For PATCH changes, the classifier correctly identified 1 out of 7 (14.3\% accuracy), misclassifying 4 as MAJOR and 2 as MINOR.

\subsubsection{5.3 Drift Score Distribution}

Drift scores ranged from 0.042 to 0.79, an 18.8× variance. Figure 3 shows scores by ground truth label:

\textbf{PATCH-labeled datasets:}
\begin{itemize}
\item Minimum: 0.082 (SNLI)
\item Maximum: 0.79 (SST2)
\item MNIST→USPS: 0.597
\end{itemize}

\textbf{MINOR-labeled datasets:}
\begin{itemize}
\item GLUE QNLI: 0.042 (lowest overall)
\item GLUE MRPC: 0.053
\item GLUE RTE: 0.057
\end{itemize}

\textbf{MAJOR-labeled datasets:}
\begin{itemize}
\item MultiNLI: 0.087
\end{itemize}

Four PATCH-labeled datasets (SNLI, CoLA, WNLI, SST2) scored above the 0.07 MAJOR threshold. The MAJOR-labeled MultiNLI dataset (0.087) scored lower than the PATCH-labeled SST2 dataset (0.79), representing a 9× difference in drift magnitude despite reversed severity labels.

\subsubsection{5.4 Per-Dataset Results}

Of 14 evaluated datasets:
\begin{itemize}
\item 4 correctly classified (28.57\%)
\item 10 misclassified (71.43\%)
\end{itemize}

Correct classifications:
\begin{itemize}
\item 1 MAJOR dataset
\item 1 MINOR dataset
\item 1 PATCH dataset
\item 1 additional dataset
\end{itemize}

Misclassifications:
\begin{itemize}
\item 3 MAJOR changes missed
\item 2 MINOR changes incorrectly classified as MAJOR
\item 5 PATCH changes misclassified (4 as MAJOR, 1 as MINOR)
\end{itemize}

\subsubsection{5.5 Comparison to Baselines}

A random classifier for three classes would achieve 33.3\% accuracy. The evaluated classifier achieved 28.57\% accuracy, 4.73 percentage points below random. The classifier's precision (25\%) and recall (25\%) for MAJOR detection both fell below the random baseline of 33\%.
