
\subsubsection{4.1 Research Questions}

\begin{itemize}
\item \textbf{RQ1:} Do benchmark items distribute meaningfully across lexical, semantic, and indeterminate strata when geometry features are computed from 50K-document corpus samples?
\item \textbf{RQ2:} Do n-gram detectors exhibit recall separation between lexical (≥ 0.80) and semantic (≤ 0.40) strata?
\item \textbf{RQ3:} Does Min-K\%++ F1 vary across corpora (variance ≥ 0.15), indicating corpus-dependent detector sensitivity?
\end{itemize}

\subsubsection{4.2 Benchmarks}

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
Benchmark & HuggingFace ID & Items & Type \\
\midrule
MMLU & `cais/mmlu` & 14,042 & NLU factual QA \\
HellaSwag & `Rowan/hellaswag` & 10,042 & NLU commonsense \\
GSM8K & `openai/gsm8k` & 1,319 & Math reasoning \\
**Total** &  & **25,403** &  \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{4.3 Corpora}

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
Corpus & HuggingFace ID & N-gram Index Size & SBERT Vecs \\
\midrule
The Pile & `monology/pile-uncopyrighted` & 37,678,937 n-grams & 50,000 \\
C4 & `allenai/c4` & 17,182,929 n-grams & 50,000 \\
FineWeb* & `HuggingFaceFW/fineweb` & 25,260,267 n-grams & 50,000 \\
\bottomrule
\end{tabular}
\end{table}

*FineWeb was substituted for RedPajama-Data-1T (\texttt{togethercomputer/RedPajama-Data-1T}). This substitution was not specified in the original hypothesis design and is noted as a limitation.

\subsubsection{4.4 Models}

\textbf{LLM models:} Pythia-6.9B (\texttt{EleutherAI/pythia-6.9b}, target) and Pythia-2.8B (reference for DC-PDD), loaded in bfloat16.

\textbf{SBERT model:} \texttt{all-MiniLM-L6-v2} via sentence-transformers. FAISS IndexFlatIP over normalized embeddings.

\textbf{Critical deviation:} In the executed experiment, the FAISS index was built over randomly-streamed documents rather than per-item top-k retrieved documents. This is the design deviation that causes stratum collapse.

\subsubsection{4.5 Hardware and Runtime}

\begin{itemize}
\item GPU: NVIDIA H100 NVL (CUDA_VISIBLE_DEVICES=1)
\item Conda environment: youra-h-e1 (Python 3.10)
\item Total runtime: 5,676 seconds (~94 minutes)
\item Coder-Validator cycles completed: 1 of 5
\end{itemize}

\subsubsection{4.6 Evaluation Metrics and Gate Targets}

\begin{table}[htbp]
\centering
\begin{tabular}{ll}
\toprule
Metric & Gate Target \\
\midrule
N-gram Recall (Lexical stratum) & ≥ 0.80 \\
N-gram Recall (Semantic stratum) & ≤ 0.40 \\
Min-K\%++ F1 Variance across corpora & ≥ 0.15 \\
Indeterminacy Rate & [0.10, 0.50] \\
\bottomrule
\end{tabular}
\end{table}

Gate type: MUST_WORK. Failure condition: stop entire verification chain; report failure as negative finding.

\subsubsection{4.7 Dry Run}

A dry run at 0.01× scale (500 corpus documents) was conducted prior to the full experiment. The dry run produced 3-way stratification: lexical = 13 items, semantic = 12 items, indeterminate = 25 items. This dry-run result was not predictive of full-scale behavior; the full run at 50K documents produced complete stratum collapse.

