
\subsubsection{5.1 Classification Accuracy}

Table 1 shows results across all feature types.

\begin{table}[htbp]
\centering
\begin{tabular}{lllllll}
\toprule
Feature Type & Features & Test Acc & Train Acc & Random Acc & Within-Family ResNet & Within-Family DenseNet \\
\midrule
Global stats & 4 & 100\% (4/4) & 93.8\% (15/16) & Not tested & Not tested & Not tested \\
Layer-wise & 6 & 100\% (4/4) & 81.3\% (13/16) & 100\% (4/4) & Not tested & Not tested \\
Architectural & 8 & 100\% (4/4) & 93.8\% (15/16) & 100\% (4/4) & 100\% & 100\% \\
Batch norm & 6 & 75\% (3/4) & 81.2\% (13/16) & 75\% (3/4) & 100\% & 100\% \\
\bottomrule
\end{tabular}
\end{table}

Global statistics, layer-wise features, and architectural features each achieved 100\% test accuracy. Batch normalization features achieved 75\% (3/4 correct), with resnet152 misclassified as shallow.

Test model predictions: alexnet (actual: shallow, predicted: shallow by all methods), vgg13 (actual: shallow, predicted: shallow by all methods), resnet152 (actual: deep, predicted: deep except by batch normalization features which predicted shallow), wide_resnet50_2 (actual: deep, predicted: deep by all methods).

Confusion matrix for the three 100\%-accurate feature types:
```
              Predicted
           Shallow  Deep
Actual  
Shallow      2       0
Deep         0       2
```

\subsubsection{5.2 Random Initialization Results}

Random initialization tests compared pretrained models to randomly initialized models with identical architectures.

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
Feature Type & Pretrained Accuracy & Random Accuracy & Difference \\
\midrule
Layer-wise & 100\% & 100\% & 0\% \\
Architectural & 100\% & 100\% & 0\% \\
Batch normalization & 75\% & 75\% & 0\% \\
\bottomrule
\end{tabular}
\end{table}

Zero performance difference indicates that discriminative features are present in network structure independent of training. Random models achieved identical classification accuracy to pretrained models for all tested feature types.

\subsubsection{5.3 Within-Family Validation}

Within-family validation tested depth classification within single architecture families.

\begin{table}[htbp]
\centering
\begin{tabular}{lllll}
\toprule
Family & Number of Models & Shallow/Deep Split & Architectural Features Accuracy & Batch Norm Features Accuracy \\
\midrule
ResNet & 9 & 2 shallow, 7 deep & 100\% & 100\% \\
DenseNet & 4 & 1 shallow, 3 deep & 100\% & 100\% \\
VGG & 4 & 4 shallow, 0 deep & Not tested (no deep variants) & Not tested (no deep variants) \\
\bottomrule
\end{tabular}
\end{table}

Both ResNet and DenseNet families achieved 100\% within-family accuracy. VGG could not be tested because all torchvision VGG models are shallow.

\subsubsection{5.4 Feature Importance}

Logistic regression coefficients for architectural features (highest magnitude indicates strongest discriminative power):

\begin{table}[htbp]
\centering
\begin{tabular}{lll}
\toprule
Feature & Coefficient & Direction \\
\midrule
Bottleneck ratio & +0.956 & Positive (higher ratio → deep) \\
Layer count & +0.932 & Positive (more layers → deep) \\
Residual blocks & +0.606 & Positive (more blocks → deep) \\
Residual norm & +0.479 & Positive \\
Dense connections & +0.396 & Positive (more connections → deep) \\
Skip connections & +0.349 & Positive \\
Architecture family & +0.349 & Positive \\
Transition layers & -0.026 & Near zero (weak discriminator) \\
\bottomrule
\end{tabular}
\end{table}

Bottleneck ratio and layer count showed the strongest associations with depth classification. Transition layer count showed minimal discriminative power.

For global statistics features:

\begin{table}[htbp]
\centering
\begin{tabular}{lll}
\toprule
Feature & Coefficient & Direction \\
\midrule
Mean norm & -0.85 & Negative (lower mean → deep) \\
Std norm & -0.35 & Negative \\
Min norm & -0.13 & Negative \\
Max norm & +0.12 & Positive \\
\bottomrule
\end{tabular}
\end{table}

Mean weight magnitude showed the strongest association, with deep models exhibiting lower mean values.

\subsubsection{5.5 Feature Distributions}

Bottleneck ratio by model:
\begin{itemize}
\item VGG models: 0.0 (no 1×1 convolutions)
\item ResNet-18: 0.08
\item ResNet-50: 0.51
\item ResNet-152: 0.68
\item DenseNet-201: 0.67
\end{itemize}

Layer count by model:
\begin{itemize}
\item AlexNet: 8 layers
\item VGG-19: 19 layers
\item ResNet-34: 34 layers
\item ResNet-50: 50 layers
\item ResNet-152: 152 layers
\item DenseNet-201: 201 layers
\end{itemize}

Residual block count by model:
\begin{itemize}
\item VGG models: 0 blocks
\item ResNet-18: 8 blocks
\item ResNet-50: 16 blocks
\item ResNet-152: 50 blocks
\end{itemize}
