
\subsubsection{4.1 Models}

\begin{table*}[htbp]
\centering
\resizebox{\dimexpr 2\columnwidth+\columnsep\relax}{!}{%
\begin{tabular}{lllll}
\toprule
Model Size & Base & SFT & DPO & PPO \\
\midrule
1.4B & EleutherAI/pythia-1.4b & lomahony/pythia-1.4b-helpful-sft & Leogrin/eleuther-pythia1.4b-hh-dpo & usvsnsp/pythia-1.4b-ppo \\
2.8B & EleutherAI/pythia-2.8b & lomahony/pythia-2.8b-helpful-sft & Leogrin/eleuther-pythia2.8b-hh-dpo & usvsnsp/pythia-2.8b-ppo \\
6.9B & EleutherAI/pythia-6.9b & lomahony/pythia-6.9b-helpful-sft & Leogrin/eleuther-pythia6.9b-hh-dpo & usvsnsp/pythia-6.9b-ppo \\
\bottomrule
\end{tabular}
}
\end{table*}

All aligned variants share the Pythia pretraining checkpoint. Public fallback checkpoints (trained on Anthropic HH preference data) were used in place of the primary RLHFlow checkpoints.

\subsubsection{4.2 Datasets}

\textbf{MMLU \citep{hendrycks2021mmlu}:} 14,042 multiple-choice questions across 57 subjects. The 4-option forced-choice format provides a 4-dimensional probability vector for Spearman rho analysis. Evaluated with 4-shot prompting.

\textbf{TruthfulQA MC1 \citep{lin2022truthfulqa}:} 817 questions with adversarially crafted plausible-but-false alternatives. Used exclusively as the H3 framing susceptibility diagnostic (0-shot).

\subsubsection{4.3 Metrics}

\begin{itemize}
\item \textbf{Primary:} $\Delta Brier$ Reliability (aligned minus base)
\item \textbf{Secondary:} $\Delta ECE$ (15-bin equal-width), Spearman rho (per-item, per pair), $\Delta margin$ (pre-softmax), argmax redistribution rate, $\Delta ECE$ ratio (TruthfulQA/MMLU)
\end{itemize}

\subsubsection{4.4 Implementation}

All evaluations used lm-eval-harness v0.4.11 with --log\_samples. Calibration analysis was implemented in Python using scipy.stats (Spearman rho), numpy (Brier decomposition, argmax partition), and a 15-bin equal-width ECE implementation following Guo et al. (2017). Bootstrap resampling used seed = 42. Calibration bins span [0, 1] in 15 equal-width intervals, consistent with Xie et al. (2024).
