
\subsubsection{A. H-E1 Per-Example Spearman ρ Summary}

\begin{table}[htbp]
\centering
\begin{tabular}{ll}
\toprule
Metric & Value \\
\midrule
Mean ρ & 0.3662 \\
Standard deviation & 0.0759 \\
Minimum ρ (example) & ~0.17 \\
Maximum ρ (example) & ~0.58 \\
Fraction ρ < 0.70 & 100\% (100/100) \\
Borderline [0.65, 0.75] & 0 examples \\
N & 100 MNLI validation_matched examples \\
Seed & 42 \\
\bottomrule
\end{tabular}
\end{table}

\textit{Source: h-e1/results/spearman_correlation_results.json.}

\subsubsection{B. H-M1 Mechanism Verification Indicators}

\begin{table}[htbp]
\centering
\begin{tabular}{lll}
\toprule
Indicator & Value & Source \\
\midrule
locret_grad_received & True & h-m1/04_validation.md \\
cis_shape_correct & True (B, L, 8) & h-m1/experiment_results.json \\
eviction_active & True (retained ratio < 0.55) & h-m1/experiment_results.json \\
accuracy_improved & True (+1.50pp vs B1) & h-m1/experiment_results.json \\
gate_result & PARTIAL (gap=1.50pp < threshold=2.0pp) & h-m1/experiment_results.json \\
Training scale & 1 seed, 1 epoch, 500 samples & h-m1/04_validation.md \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{C. H-M2 Per-Seed LongBench F1 — PoC Model}

The following values are from the tiny PoC model (d=64, 2-layer transformer). The model outputs class-index tokens that do not match QA answer strings; F1 = 0.000 for all conditions is the correct measured value at this model scale and reflects the model's inability to generate coherent text, not a property of joint training.

\begin{table}[htbp]
\centering
\begin{tabular}{lll}
\toprule
Task & JointLoRA-KV & B3 Baseline \\
\midrule
NarrativeQA & 0.000 & 0.000 \\
Qasper & 0.000 & 0.000 \\
MultiFieldQA-en & 0.000 & 0.000 \\
Mean F1 & 0.000 & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\textit{Source: h-m2/results/aggregated_results.json.}

The gate criterion (joint F1 ≥ B3 F1) is satisfied by equality. These values do not reflect task performance on a real model.

\subsubsection{D. Discrepancy Note on H-M2 F1 Values}

The verification_state.yaml and the 045_validated_hypothesis.md synthesis document report "joint_mean_f1=0.3375 ≥ b3_mean_f1=0.3354" as key findings from H-M2. The aggregated_results.json in h-m2/results/ shows F1 = 0.000 for all conditions. The 04_validation.md for H-M2 explicitly states that the PoC model's F1 = 0.000 reflects the tiny-model evaluation, while the experiment log appendix records "Joint mean F1: 0.3375, B3 mean F1: 0.3354." The source of the 0.3375/0.3354 values is not traceable to the raw output file; these values may reflect an intermediate computation or a different evaluation run not reflected in the final aggregated results file. This paper reports the values directly from aggregated_results.json (F1 = 0.000) as the primary source, and flags the discrepancy.
