
\subsubsection{4.1 Research Questions}

\begin{itemize}
\item \textbf{RQ1:} Does fastText quality filtering create systematic, monotonic changes in H(occupation|demographic)?
\item \textbf{RQ2:} Do conditional log-odds of demographic-occupation co-occurrences correlate with filtering intensity?
\item \textbf{RQ3:} Does a model trained on filtered corpora represent corpus demographic structure in its logit space?
\end{itemize}

\subsubsection{4.2 Dataset}

DCLM-POOL (mlfoundations/dclm-baseline-1.0): CommonCrawl-derived text used as the standard substrate for fastText quality filtering. Streaming access was used with approximately 50,000-document subsets per configuration (8 configurations, approximately 400,000 documents total).

\subsubsection{4.3 Implementation}

\textbf{Corpus audit (H-E1, H-M1):} Python 3.10, HuggingFace datasets (streaming), numpy, scipy.stats. Pipeline: CorpusFilter, EntropyMeasure, LogOddsComputer, StatisticalTests. Unit tests: 57/57 passing for H-E1; 26/26 tasks completed for H-M1.

\textbf{Model training (H-M2):} Proxy model trained with hf\_trainer\_fallback. Learning rate 2e-5, batch size 256, cross-entropy loss. Approximately 95,368 training steps per configuration. N = 2,160 probe samples per configuration (50+ templates x 20 occupation pairs x 2 pronouns).

\subsubsection{4.4 Evaluation Metrics}

\begin{table*}[htbp]
\centering
\resizebox{\dimexpr 2\columnwidth+\columnsep\relax}{!}{%
\begin{tabular}{lllll}
\toprule
Metric & Gate Criterion & Applies To &  &  \\
\midrule
H(occ\ & demo) relative change & >= 5\% & H-E1 &  \\
Spearman rho (entropy vs. filter intensity) & rho != 0, p < 0.05 & H-E1 &  &  \\
Bootstrap CI for H(C5) - H(C1) & Excludes zero & H-E1 &  &  \\
Mean log-odds Spearman rho & \ & rho\ & > 0, p < 0.05 & H-M1 \\
Logit margin Spearman rho & rho > 0, p < 0.01 & H-M2 &  &  \\
Negative control \ & C7 - C0\ &  & Detectable difference & H-M2 \\
\bottomrule
\end{tabular}
}
\end{table*}

