
Low-Rank Adaptation (LoRA) enables fine-tuning of large language models using rank-8 to rank-64 matrix decompositions, reducing trainable parameters by orders of magnitude while maintaining performance comparable to full fine-tuning. This empirical success has led to an implicit assumption in portions of the model compression literature: if low-rank updates $\Delta W$ work effectively, perhaps pre-trained weights W themselves exhibit low-rank structure.

We tested this assumption directly through singular value decomposition analysis of projection weight matrices in Mistral-7B, a 32-layer decoder-only Transformer with 7B parameters. The analysis targeted deep layers (L $\geq$ 20) where semantic compression theories predict the strongest low-rank signatures. Results show effective ranks of 1554-1647 at 99\% variance threshold, approaching ~40\% of the model dimension (4096). This represents 6-$7\times$ higher than thresholds commonly assumed for certain post-hoc compression approaches and $50\times$ higher than typical LoRA adaptation ranks.

\subsubsection{Scope and Objectives}

This work measures a specific structural property: the effective rank of learned projection weight matrices in pre-trained models. The measurement addresses whether assumptions derived from LoRA's success---namely, that pre-trained weights might be inherently low-rank---hold empirically at the 7B parameter scale.

We do not propose new methods, compare competing techniques, or evaluate model performance. This is a measurement study establishing empirical ground truth for a structural property relevant to compression research.

\subsubsection{Contributions}

\begin{enumerate}
\item Direct measurement of effective rank in pre-trained 7B-scale Transformer projection weights, revealing near full-rank structure (r\_eff ~ 1600)
\item Empirical refutation of the assumption that projection weight matrices exhibit low-rank structure at this scale
\item Validated SVD-based analysis methodology applicable to other architectural families and scales
\item Clarification that LoRA's success with low-rank updates (r ~ 8-64) is independent from the rank structure of pre-trained weights (r\_eff ~ 1600)
\end{enumerate}
