
\subsubsection{Effective Rank Computation}

For a weight matrix W $\in \mathbb{R}^(d_out \times$ d\_in), the singular value decomposition is W = $U\Sigma V^T$. The effective rank at variance threshold $\tau$ is defined as:

$r_eff(W, \tau ) = min{r : \Sigma _{i=1}^r \sigma _{i}^2 / \Sigma _{i=1}^{rank(W)} \sigma _{i}^2 \geq \tau }$

We use $\tau$ = 0.99 (99\% variance threshold). This quantifies how many singular values encode 99\% of the matrix information. We chose r\_eff < 256 as the test threshold based on state size constraints in selective SSM architectures (N $\leq$ 1024) and assumptions in post-hoc conversion literature.

\subsubsection{Target Matrices and Layers}

We analyzed Query (Q), Key (K), and Value (V) projection weight matrices in multi-head attention layers. For Mistral-7B with hidden dimension 4096 and 32 attention heads, each projection is $W_{Q,K,V} \in \mathbb{R}^(4096 \times$ 4096). These are learned parameters stored in the model, not runtime-computed attention patterns.

Analysis focused on layers 20-31 (deep layers where L $\geq$ 20 in the 32-layer model). Semantic compression theories predict deep layers should exhibit the strongest low-rank signatures.

\subsubsection{Operator Entropy Measurement}

We measured operator entropy via log-determinant of covariance matrices:

H(W) = log det((1/n) $W^T$ W)

The hypothesis predicted monotonically decreasing entropy with layer depth. We tested this via linear regression H(L) = $\beta \times$ L + intercept with null hypothesis $H_0$: $\beta \geq$ 0 and alternative $H_A$: $\beta$ < 0, p < 0.01.

\subsubsection{Model and Computational Setup}

Analysis used Mistral-7B-v0.1, a 32-layer decoder-only Transformer with 7B parameters, 4096 hidden dimension, and 32 attention heads. Mistral shares architectural characteristics with the LLaMA family (32 layers, similar hidden sizes).

SVD computation for $4096\times 4096$ matrices used NumPy's \texttt{np.linalg.svd} with float64 precision for numerical accuracy. Analysis was conducted on NVIDIA H100 GPU for weight loading; SVD computation is CPU-bound.

\subsubsection{Two-Phase Validation}

To ensure measurement reliability separates from hypothesis outcomes:

\textbf{Phase 1 (h-e1): Methodology Validation}
\begin{itemize}
\item Implement SVD-based rank computation pipeline
\item Test on Mistral-7B with reduced samples
\item Verify numerical stability
\item Gate: Methodology must pass before hypothesis testing
\end{itemize}

\textbf{Phase 2 (h-m1): Hypothesis Testing}
\begin{itemize}
\item Apply validated methodology to target layers 20-31
\item Measure r\_eff and entropy
\item Evaluate against thresholds
\end{itemize}

This separation ensures that hypothesis refutation reflects genuine properties of the weights, not implementation errors.
