
\subsubsection{3.1 Problem Formulation}

Let Z = ${(w_i, g_i)}_{i=1}^N$ be a model zoo, where w\_i are the weight parameters of the i-th trained network and g\_i = train\_loss\_i - test\_loss\_i is the generalization gap. An encoder phi: $R^D$ -> $R^d$ followed by a regression head h: $R^d$ -> R predicts g from w.

A permutation-equivariant encoder satisfies phi($w^pi$) = pi-hat(phi(w)) for a corresponding action pi-hat, such that any permutation-invariant head h produces h(phi($w^pi$)) = h(phi(w)).

Encoders are evaluated at permutation severity s in {0, 0.25, 0.5, 1.0}, where s is the fraction of neurons randomly permuted within each layer at test time. The primary sensitivity metric is delta-rho = rho(s=0) - rho(s=1.0), where rho is the Spearman rank correlation between predicted and true generalization gap.

\subsubsection{3.2 Encoder Suite}

Six encoder variants were compared:

\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{llll}
\toprule
Encoder & Architecture & Parameters & Permutation Handling \\
\midrule
flat-MLP & 3-layer MLP, hidden=512 & 3.04M & None \\
flat-MLP+aug & Same as flat-MLP & 3.04M & Training-time augmentation \\
flat-MLP+canon & Same, L2-normed inputs & 3.04M & L2 canonicalization \\
NFT-base & NFT, d\_model=128, n\_heads=4 & 75K & Architectural equivariance \\
NFT+aug & NFT + augmentation & 75K & Architectural + augmentation \\
Oracle-canon & flat-MLP, optimal alignment & 3.04M & Oracle (upper bound) \\
\bottomrule
\end{tabular}
}
\end{table}

The flat-MLP encoder concatenates all weight matrices into a single 4,912-dimensional vector and applies a 3-layer MLP with hidden dimension 512. The NFT encoder represents each neuron's incoming weight vector (fan\_in = 16) as a token, projects it to d\_model = 128 dimensions, and applies multi-head attention (n\_heads = 4) within each layer. Cross-layer aggregation combines per-layer embeddings into a fixed-size representation. The flat-MLP+aug encoder applies random neuron permutations to input weight vectors during training, following Schurholt et al. (2021). The flat-MLP+canon encoder normalizes each neuron's weight vector to unit L2 norm before encoding. Oracle canonicalization aligns each test model's neurons to a reference model via the optimal permutation minimizing L2 distance.

\subsubsection{3.3 Mediation Analysis}

Following the hierarchical regression framework of Baron and Kenny (1986), generalization gap was regressed on flat-MLP+aug embeddings (yielding R-squared\_aug) and on NFT-base embeddings (yielding R-squared\_NFT). The mediation effect is delta-R-squared = R-squared\_NFT - R-squared\_aug. A pre-specified gate condition required delta-R-squared >= 0.10.

\subsubsection{3.4 Statistical Protocol}

Bootstrap testing for delta-rho significance used n = 10,000 paired resamples from the test set, with Holm-Bonferroni correction for multiple comparisons across severity levels. Two-sided p-values tested the null hypothesis delta-rho = 0.
