
\subsubsection{Controlled Adapter Generation}

Adapters are generated in-house with verified provenance. A single base model checkpoint (TinyLlama-1.1B-Chat-v1.0) is fine-tuned with identical LoRA configuration across all tasks. The LoRA configuration uses rank r = 32, alpha = 64, dropout = 0.05, targeting all attention and MLP projections (q\_proj, k\_proj, v\_proj, o\_proj, up\_proj, down\_proj, gate\_proj).

All adapters are trained with identical hyperparameters: learning rate 2 $\times 10^{-4}$, batch size 8, 3 epochs, AdamW optimizer, warmup ratio 0.1, and maximum sequence length 512. For each task, 5 adapters are trained with different random seeds (42, 43, 44, 45, 46) to assess within-task variance.

\subsubsection{Grassmann Distance Computation}

Given two B matrices $B_1, B_2 \in \mathbb{R}^{d\times r}$, the Grassmann distance is computed as follows:

\begin{enumerate}
\item Compute orthonormal bases for the column spaces via SVD
\item Compute principal angles $\theta_1, \ldots, \theta_r$ between the subspaces
\item Grassmann distance: $d_G(B_1, B_2) = \sqrt{\sum_j \theta_j^{2}}$
\end{enumerate}

This distance is zero when subspaces are identical and maximal when subspaces are orthogonal.

\subsubsection{Statistical Analysis}

The main hypothesis is tested using the Mann-Whitney U test comparing within-category vs. between-category distance distributions. Cohen's d is reported as effect size. Spearman rank correlation validates the relationship between geometric distance and semantic similarity. Bootstrap confidence intervals (1000 iterations) are computed for all primary metrics.
