\section{Experimental Setup}
\label{sec:experiments}

We design experiments to answer three research questions:

\textbf{RQ1:} Do stylistic preference coefficients ($\beta_L$, $\beta_H$, $\beta_S$) shift directionally across HH-RLHF annotation strata after controlling for $Q_\text{early}$?

\textbf{RQ2:} Is the annotation preference gradient in WebGPT aligned with the AI-typicality direction in sentence embedding space, and is this alignment discriminant-valid?

\textbf{RQ3:} What minimum data requirements are needed to confirm individual annotator causal adaptation?

\subsection{Datasets}

\begin{table}[h]
\centering
\caption{Dataset summary.}
\begin{tabular}{lllll}
\toprule
Dataset & Rows & Strata & Rows/Stratum & Task \\
\midrule
HH-RLHF & 160,800 & 3 (index-based) & 53,600 & Helpful/Harmless preference \\
WebGPT & 19,578 & Between-group tercile & --- & Summary preference \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Baselines}

\textbf{Random temporal split:} Verifies that temporal round effects are not sampling artifacts.

\textbf{$Q_\text{early}$ quality predictor (frozen):} Isolates stylistic preference from semantic quality.

\textbf{Placebo AI-typicality vector:} A randomly oriented unit vector in the same embedding space, used to validate measurement specificity (discriminant validity).

\subsection{Evaluation Metrics}

\begin{table}[h]
\centering
\caption{Pre-registered evaluation metrics and gate thresholds.}
\begin{tabular}{llll}
\toprule
Metric & Definition & Gate Threshold & Hypothesis \\
\midrule
$n\_\text{directional}$ & Features with non-overlapping 95\% bootstrap CIs & $\geq 2$ of 3 & H-M2 \\
sign\_consistent & All stylistic $\Delta\beta > 0$ & true & H-M2 \\
$\beta_\text{exposure}$ & Exposure tercile regression coefficient & $> 0$, $p < 0.05$ & H-M1 \\
Placebo $p$-value & Empirical $p$ for random vector projection & $> 0.05$ & H-M1 \\
Interaction $p$ & Round $\times$ ambiguity interaction & $< 0.0167$ (Bonferroni) & H-E1 \\
\bottomrule
\end{tabular}
\end{table}

Statistical significance is evaluated at $\alpha = 0.05$ (unadjusted) and $\alpha = 0.0167$ (Bonferroni-corrected for $k = 3$ stylistic features). Bootstrap confidence intervals use 2,000 stratified resamples.

\subsection{Implementation Details}

\textbf{Feature engineering:} Stylistic features extracted from preferred/rejected response text differences; standardized with a shared \texttt{StandardScaler} fit on round-1 training data.

\textbf{Logistic regression:} scikit-learn \texttt{LogisticRegression} with $C = 1.0$, solver = \texttt{lbfgs}, max\_iter = 1000, class\_weight = balanced, random\_state = 42. Train/test split: 75\%/25\% stratified by stratum.

\textbf{Bootstrap CI:} 2,000 stratified resamples per stratum (H-M2); 200 resamples (H-E1).

\textbf{Sentence encoder:} \texttt{all-MiniLM-L6-v2}~\citep{Reimers2019SentenceBERT}, frozen weights, batch encoding at inference (H-M1).

\textbf{Compute:} Single NVIDIA H100 NVL GPU (\texttt{CUDA\_VISIBLE\_DEVICES=0}). H-E1 runtime: $\sim$49 minutes. H-M1 runtime: $\sim$8 minutes. H-M2 runtime: $\sim$2.5 minutes.

\textbf{Reproducibility:} All random states fixed (random\_state = 42). Code available with 26 unit tests (all passing).
