
\subsubsection{6.1 Interpretation}

The 13\% relative improvement from semantic clustering (0.78 vs 0.69 AUROC) demonstrates that grouping semantically equivalent answers captures uncertainty more effectively than exact string matching when computational budget is matched. The ablation design attributes this improvement to the clustering mechanism.

Low correlations (maximum 0.21 excluding the implementation issue) indicate that methods measure distinct aspects of uncertainty. Semantic entropy captures semantic diversity, self-consistency measures sampling agreement, and verbalized confidence reflects introspection. These orthogonal signals could be combined for robust uncertainty estimation.

The null result on error-type signatures reveals that dataset-level partitioning is insufficient. TruthfulQA questions may trigger multiple competing misconceptions (producing high diversity) rather than single confident errors (low diversity). Dataset labels are too coarse-grained for error-type analysis. Instance-level features would be required.

\subsubsection{6.2 Limitations}

\textbf{Pilot Scale:} With 100 samples per dataset, results demonstrate proof-of-concept. Confidence intervals for AUROC estimates are wider than in full-scale studies. Scaling to complete test sets would strengthen claims.

\textbf{Single Model:} Results are specific to Mistral-7B and may not generalize across model families or scales. Multi-model validation would clarify whether findings are universal properties.

\textbf{Implementation Issue:} Token variance collapsed with self-consistency due to implementation bug, preventing independent evaluation of distributional sharpness. Reimplementation with logit-level analysis is required.

\textbf{Dataset-Level Partitioning:} The error-type hypothesis assumed dataset choice reflects error properties. The null result reveals this assumption is false and motivates instance-level analysis.

\textbf{No Calibration Analysis:} We focused on discrimination (AUROC) rather than calibration (Expected Calibration Error). Calibration analysis remains future work.

\subsubsection{6.3 Implications}

Semantic entropy demonstrates measurable improvement in controlled experiments for factual question answering on knowledge-gap errors using Mistral-7B. Methods measure orthogonal dimensions, justifying multi-method approaches for robust uncertainty quantification. Dataset labels are insufficient for error-type characterization; instance-level features are required.
