
\subsubsection{5.1 Semantic Clustering Contribution}

Semantic entropy achieved AUROC 0.78 on NaturalQuestions unanswerable questions, compared to 0.69 for the ensemble baseline. The difference of 0.09 exceeds the threshold of 0.07 and represents a 13\% relative improvement. Both methods used identical K=10 samples, isolating the clustering mechanism. The semantic entropy absolute AUROC of 0.78 exceeds the minimum threshold of 0.70.

\subsubsection{5.2 Method Independence}

Correlation matrix:

\begin{table}[htbp]
\centering
\begin{tabular}{lllll}
\toprule
 & Semantic Entropy & Self-Consistency & Token Variance & Verbalized Conf \\
\midrule
Semantic Entropy & 1.000 & -0.022 & -0.022 & 0.208 \\
Self-Consistency & -0.022 & 1.000 & 1.000 & 0.020 \\
Token Variance & -0.022 & 1.000 & 1.000 & 0.020 \\
Verbalized Conf & 0.208 & 0.020 & 0.020 & 1.000 \\
\bottomrule
\end{tabular}
\end{table}

Self-consistency and token variance show perfect correlation (1.0), indicating an implementation issue where both methods computed the same statistic. Post-validation analysis identified that both used agreement rate calculation. Token variance implementation requires correction with logit-level probability analysis.

Excluding the implementation bug, all pairwise correlations fall between -0.022 and 0.208. Maximum correlation 0.21 is well below the 0.7 threshold, indicating that semantic entropy, self-consistency, and verbalized confidence measure distinct uncertainty dimensions.

\subsubsection{5.3 Error Type Signatures}

Diversity comparison:
\begin{itemize}
\item NaturalQuestions mean diversity: 0.975 (SD 0.488)
\item TruthfulQA mean diversity: 1.077 (SD 0.515)
\item Difference: 0.102 (opposite predicted direction)
\item Statistical test: t = -1.418, p = 0.158 (not significant)
\end{itemize}

Agreement comparison:
\begin{itemize}
\item NaturalQuestions mean agreement: 0.200 (SD 0.000)
\item TruthfulQA mean agreement: 0.204 (SD 0.028)
\item Difference: 0.004, t = -1.421, p = 0.157 (not significant)
\end{itemize}

The difference is neither statistically significant (p > 0.05) nor in the predicted direction. Dataset-level partitioning does not cleanly separate error types by uncertainty signature.
