
This work presents a controlled benchmark comparing flat MLP, Deep Sets, and NFN weight-space encoders at matched approximately 500K parameters on the Schurholt MNIST-CNN model zoo. The primary result is Δρ(NFN − flat MLP) = 0.5119 [95\% CI: 0.381, 0.638], computed against an untrained flat MLP baseline. A monotone ordering ρ(flat MLP) = 0.169 < ρ(Deep Sets) = 0.447 < ρ(NFN) = 0.681 is confirmed. Permutation sensitivity measurement reveals that the flat MLP assigns substantially different embeddings to equivalent weight configurations (sensitivity = 0.649), while the NFN collapses them to near-zero (sensitivity = 7.34 × 10⁻⁷, approximately 885,000× lower).

The results are subject to several limitations: all claims apply only to the MNIST-CNN zoo (CIFAR-10 was unavailable); the flat MLP uses an architecturally constrained single-layer configuration that may inflate Δρ; only one training seed was used; and the orbit existence result was established on a different dataset variant than the encoder experiments.

An unexpected finding — that NFN advantage peaks for low-accuracy models (ρ = 0.856) and inverts for high-accuracy models (ρ = −0.314) — was not anticipated by the pre-registered predictions. This pattern was not characterized with tier-level bootstrap CIs and its generality is unknown.

Future work should address: (1) CIFAR-10 cross-zoo validation; (2) multi-layer flat MLP baselines at the same parameter budget; (3) capacity curve analysis across 50K–2M parameters; (4) multi-seed training replication; and (5) direct orbit existence measurement on the hyp_rand zoo variant.

