
\subsubsection{6.1 Interpretation of Main Results}

The four experiments form a sequential chain of evidence. The seed-only zoo contains universally permutation-distinct weight configurations (orbit_proportion = 1.000). Flat MLPs at 500K parameters cannot learn permutation invariance from training data (sensitivity = 0.649). NFN encoders eliminate it by construction (sensitivity = 7.34 × 10⁻⁷). This 885,000× reduction in sensitivity is associated with a 4× improvement in Spearman ρ (0.169 → 0.681).

The association between permutation sensitivity reduction and ρ improvement is consistent with the capacity reallocation hypothesis: encoder parameters freed from orbit disambiguation are available for accuracy-predictive feature learning. However, the NFN and flat MLP encoders differ in multiple architectural respects beyond equivariance, including weight-sharing patterns and optimization landscape properties. The observed ρ improvement cannot be attributed solely to equivariance; confounds cannot be ruled out from this observational benchmark.

The monotone symmetry spectrum (flat MLP < Deep Sets < NFN) indicates that symmetry exploitation is a continuous design axis: invariance (Deep Sets) provides approximately half the total equivariance advantage (Δρ ≈ 0.28 of the total 0.51), while equivariance (NFN) provides a further increment (Δρ ≈ 0.23 over Deep Sets).

\subsubsection{6.2 Limitations}

\textbf{L1 — Single zoo [HIGH].} The CIFAR-10 experiment was not executed due to a data download failure. All results apply only to the Schurholt MNIST-CNN hyp_rand zoo. Cross-zoo generalization is not established.

\textbf{L2 — Flat MLP architectural bottleneck [HIGH].} Capacity matching at 500K parameters forces the flat MLP into a single hidden layer of width 193 for a 2,464-dimensional input. This is a severe architectural bottleneck that is unlikely to represent a well-tuned flat MLP configuration. The literature reports flat MLP ρ ≈ 0.5–0.7 on comparable zoos (Unterthiner et al., 2020), whereas the trained flat MLP here achieves ρ = 0.1041. The observed Δρ = 0.512 is therefore an upper bound on the gap against a capacity-matched flat MLP that uses a multi-layer architecture (e.g., hidden_dims = [512, 256]) with the same total parameter budget.

\textbf{L3 — Single training seed [LOW].} All trained encoders use seed 42 with no multi-seed replication. Training variance is uncharacterized.

\textbf{L4 — Single capacity point [MEDIUM].} The symmetry hierarchy is established at approximately 500K parameters. Whether it holds uniformly at other capacity scales (e.g., 50K–2M) is unknown.

\textbf{L5 — Dataset variant mismatch between H-E1 and H-M1/M2/M3 [MEDIUM].} The orbit existence analysis (H-E1) used the seed-only variant (976 checkpoints, Conv(8) architecture), while all encoder experiments used the hyp_rand variant (2,249 checkpoints, Conv(32)-Conv(64) architecture). The causal chain (orbit existence → capacity waste → performance gap) bridges two architecturally distinct dataset variants. Permutation symmetry is a structural property shared by both feedforward architectures, and the orbit existence result is expected to generalize, but it was not directly confirmed on the hyp_rand variant.

\textbf{L6 — NFN ρ below Navon et al. [MEDIUM].} The capacity-matched NFN achieves ρ = 0.6806, below the ρ ≈ 0.73 reported by Navon et al. [2023]. This gap is expected given that the present NFN uses a capacity-constrained configuration (521,953 params) rather than an unconstrained architecture, a fixed training seed without architecture-specific hyperparameter tuning, and potentially different zoo splits or subsets.

\subsubsection{6.3 Accuracy-Tier Dependence}

The refutation of P3 (mid-tier dominance) reveals an unexpected pattern: NFN advantage is concentrated among low-accuracy models (ρ = 0.856) and reverses among high-accuracy models (ρ = −0.314). One interpretation is that low-accuracy models span diverse failure modes and exhibit high weight-space diversity, providing strong equivariant signal, while high-accuracy models converge to similar near-optimal solutions where fine-grained ranking requires signals not captured by the current 500K-parameter equivariant architecture.

The negative ρ for high-accuracy models means the current NFN configuration is not suitable for model selection in settings where distinguishing the best-performing models is the goal. This represents a practical limitation not evident from the aggregate Δρ = 0.512 result.

Tier-level bootstrap CIs were not computed due to small per-tier sample size (n ≈ 112); the tier-level point estimates should be interpreted accordingly.

