
\subsubsection{3.1 Dual-Metric Heterogeneity Framework}

We define two independent metrics for characterizing solution diversity:

\textbf{Normalized Hamming Distance (d/n):} For a SAT instance with n variables, ground truth assignment x*, and generated assignment x̂, the normalized Hamming distance is:

d/n = (1/n) Σᵢ 𝟙[x̂ᵢ ≠ x*ᵢ]

This measures the proportion of variables that differ from the ground truth, ranging from 0 (perfect match) to 1 (complete disagreement).

\textbf{Violation Entropy (H):} For an instance with m clauses, let pⱼ be the proportion of total violations occurring in clause j. The violation entropy is:

H = -Σⱼ pⱼ log₂ pⱼ

This measures how violations are distributed across the constraint space. Higher entropy indicates diffuse violations across many clauses; lower entropy indicates concentrated violations.

\textbf{Dual-Metric Gate:} We test whether an architecture generates heterogeneous near-solutions by requiring:
\begin{itemize}
\item d/n range > 0.20 (solutions span at least 20 percentage points in distance)
\item Entropy range > 2.0 (violations exhibit at least 2.0 bits of structural variation)
\end{itemize}

Both criteria must be satisfied for the architecture to be considered capable of generating diverse basin entry conditions.

\subsubsection{3.2 Baseline Architecture}

We implement baseline NeuroSAT without modifications. The architecture operates on a bipartite literal-clause graph with the following structure:

\begin{itemize}
\item Hidden dimension: 128
\item Message-passing rounds: 32
\item LSTM-based state updates for literals and clauses
\item Message MLPs: 3 hidden layers with ReLU activation
\item Initialization: Single learned 128-dimensional parameter vector shared across all instances
\item No stochastic sampling, ensemble, or explicit diversity mechanisms
\end{itemize}

\textbf{Training Objective:} Unsupervised loss for SAT/UNSAT classification:
L = -log p(label | final embeddings)

where label is 1 for satisfiable instances and 0 for unsatisfiable instances.

\subsubsection{3.3 Dataset and Evaluation}

\textbf{Dataset:} G4SATBench 3-SAT Easy difficulty (10-40 variables, clause-to-variable ratio approximately 4.2-4.3).

\textbf{Training:} 20 instances (proof-of-concept scale; full dataset contains 80,000 instances). Adam optimizer with learning rate 10⁻⁴, ReduceLROnPlateau scheduler, early stopping with patience of 20 epochs.

\textbf{Evaluation:} 8 satisfiable instances from the test set (2 unsatisfiable instances filtered because d/n and entropy are undefined without ground truth). For each instance:
\begin{enumerate}
\item Generate assignment via trained NeuroSAT
\item Compute d/n against ground truth
\item Compute violation entropy H from clause violation distribution
\item Measure distribution statistics (mean, standard deviation, quartiles, range)
\end{enumerate}
