
\subsubsection{5.1 Gate Results}

The dual-metric gate produced an asymmetric outcome:

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
Metric & Observed & Threshold & Result \\
\midrule
d/n range & 0.265 & > 0.20 & PASS \\
Entropy range & 1.145 & > 2.0 & FAIL \\
Combined & 1/2 criteria & 2/2 required & FAIL \\
\bottomrule
\end{tabular}
\end{table}

The d/n range of 0.265 exceeds the 0.20 threshold by 32\%, indicating that the architecture generates solutions spanning different distances from ground truth. The entropy range of 1.145 falls below the 2.0 threshold by 43\%, indicating that violations are distributed similarly across instances despite varying distances.

\subsubsection{5.2 Distribution Analysis}

\textbf{Normalized Hamming Distance:}
\begin{itemize}
\item Mean: 0.516 ± 0.089
\item Quartiles: Q1 = 0.450, Q2 = 0.526, Q3 = 0.590
\item Range (Q3 - Q1): 0.265
\item Interpretation: Solutions vary from 45\% to 59\% variable disagreement
\end{itemize}

\textbf{Violation Entropy:}
\begin{itemize}
\item Mean: 2.692 ± 0.332
\item Quartiles: Q1 = 2.565, Q2 = 2.771, Q3 = 2.874
\item Range (Q3 - Q1): 1.145
\item Interpretation: Individual solutions have moderate entropy (mean 2.692), but the range across instances (1.145) is narrow
\end{itemize}

The correlation between d/n and entropy was weak (Pearson r = 0.28), confirming that these metrics capture independent dimensions of solution diversity.

\subsubsection{5.3 Training Convergence}

Training converged to loss 0.693 at epoch 32, matching the theoretical minimum for balanced binary classification. The model fully optimized its training objective (SAT/UNSAT classification), yet structural diversity remained limited. This observation suggests that the architectural design, rather than optimization quality, constrains structural diversity.

\subsubsection{5.4 Per-Instance Results}

The 8 test instances ranged from 12 to 35 variables and 50 to 147 clauses. The d/n values ranged from 0.417 to 0.600 (difference: 0.183), while entropy values ranged from 2.565 to 2.874 (difference: 0.309). Larger instances did not produce substantially different entropy values compared to smaller instances.
