
\subsubsection{5.1 Per-Model Effect Sizes}

Table 1 presents Cohen's d for each reward model.

\textbf{Table 1: Enumeration Effect Sizes by Reward Model}

\begin{table*}[htbp]
\centering
\resizebox{\dimexpr 2\columnwidth+\columnsep\relax}{!}{%
\begin{tabular}{llllll}
\toprule
Reward Model & Cohen's d & 95\% CI & t-statistic & p-value & n \\
\midrule
ArmoRM & 1.446 & [1.267, 1.626] & 17.81 & 6.$89\times 10^{-49}$ & 300 \\
UltraRM & 0.881 & [0.714, 1.049] & 10.36 & 1.$06\times 10^{-21}$ & 300 \\
Starling-RM & 0.378 & [0.216, 0.539] & 4.83 & 2.$22\times 10^{-6}$ & 300 \\
PairRM & 0.077 & [-0.083, 0.237] & 0.94 & 0.346 & 300 \\
\bottomrule
\end{tabular}
}
\end{table*}

Three of four models (ArmoRM, UltraRM, Starling-RM) exceeded the d = 0.3 threshold. All four models showed positive effect direction (d > 0), though only three reached statistical significance (p < 0.05).

\subsubsection{5.2 Mean Scores by Structure}

\textbf{Table 2: Mean Normalized Scores by Structure}

\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{llll}
\toprule
Reward Model & Mean (Enumerated) & Mean (Synthesized) & Difference \\
\midrule
ArmoRM & 0.646 & 0.397 & +0.250 \\
UltraRM & 0.639 & 0.466 & +0.173 \\
Starling-RM & 0.585 & 0.504 & +0.080 \\
PairRM & 0.538 & 0.527 & +0.011 \\
\bottomrule
\end{tabular}
}
\end{table}

\subsubsection{5.3 Gate Condition Evaluation}

The pre-registered primary criterion (d $\geq$ 0.3 in $\geq 2$ architecturally distinct models) was satisfied: three models exceeded the threshold.

The secondary criterion ($\geq 75$\% showing positive effect) was satisfied: all four models showed d > 0.

\subsubsection{5.4 Aggregate Statistics}

\textbf{Table 3: Pooled Statistics}

\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{ll}
\toprule
Metric & Value \\
\midrule
Pooled Cohen's d & 0.696 \\
Pooled SE & 0.115 \\
95\% CI & [0.471, 0.921] \\
$I^2$ (heterogeneity) & 99.1\% \\
Cochran's Q & 332.4 (p < 0.001) \\
\bottomrule
\end{tabular}
}
\end{table}

The high heterogeneity ($I^2$ = 99.1\%) indicates substantial variation in effect sizes across models.

\subsubsection{5.5 PairRM Results}

PairRM, the only encoder-based model tested, showed no statistically significant enumeration preference (d = 0.077, 95\% CI [-0.083, 0.237], p = 0.346). The confidence interval includes zero, indicating the effect cannot be distinguished from null.

\subsubsection{5.6 Factorial Interaction Analysis}

Analysis of structure $\times$ correctness $\times$ completeness interactions showed the enumeration effect persisted across all factorial combinations without significant interaction effects. This suggests the structural preference operates independently of the correctness and completeness manipulations.
