
\subsubsection{5.1 H-E1: Channel Permutation is an Exact Symmetry}

\textbf{Table 1: H-E1 Gate Metrics}

\begin{table}[htbp]
\centering
\begin{tabular}{lllllll}
\toprule
Architecture & Checkpoints & Permutation Runs & Mean & Delta acc &  & Orbit-PE Success Rate \\
\midrule
CNN Zoo (CIFAR-10-GS) & 200 & 2,000 & 0.000000 & 1.0 &  &  \\
Transformer Zoo (MNIST) & 250 & 2,500 & 0.000000 & 1.0 &  &  \\
Total & 450 & 4,500 & 0.000000 & 1.0 &  &  \\
\bottomrule
\end{tabular}
\end{table}

The mean |Delta acc| = 0.000000 satisfies the gate threshold of < 0.001 by several orders of magnitude. Standard deviation and maximum are also 0.000000 across all runs. Per-seed stability (Figure 8) confirms zero variance across all 10 permutation seeds. Orbit-PE computation succeeded for all layer types in both zoos using a single unified codebase (Table in Figure 6; layer coverage: 4/4 Conv2d layers, 11/11 Linear layers, 8/8 MultiheadAttention layers in the sampled checkpoints). \textbf{Gate result: PASS.}

The |Delta acc| = 0.000000 result is consistent with the theoretical proof that the input/output channel permutation group is a symmetry of any linear operator [Zhou et al., 2023; Tran-Viet et al., 2024]. The result required careful implementation: the first layer's input and the final layer's output must use identity permutations, and the Conv-to-Linear flatten transition requires expanding channel permutations to grouped indices.

\subsubsection{5.2 H-M1: Orbit-PE Computability}

\textbf{Table 2: H-M1 Gate Metrics}

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
Metric & Result & Threshold & Status \\
\midrule
overhead_ratio_mean & 1.167x & <= 1.20x & PASS \\
overhead_ratio_std & 0.061 & — & — \\
computability_rate & 1.0 (200/200) & = 1.0 & PASS \\
HAS_ARCH_BRANCHES & False & = False & PASS \\
dim_consistent & True & = True & PASS \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Table 3: Overhead by Layer Type}

\begin{table}[htbp]
\centering
\begin{tabular}{ll}
\toprule
Layer Type & Overhead Ratio (mean) \\
\midrule
Conv2d & 1.167x \\
Linear (FC) & 1.167x \\
MultiheadAttention & 1.126x \\
\bottomrule
\end{tabular}
\end{table}

Overhead is consistent across layer types (Figure 7). All 200 checkpoints computed orbit-PE successfully. \textbf{Gate result: PASS.}

\subsubsection{5.3 H-M2: Variance Decomposition}

\textbf{Table 4: H-M2 Primary Results (n = 1,000 models x 50 epochs = 50,000 checkpoints)}

\begin{table}[htbp]
\centering
\begin{tabular}{lllll}
\toprule
Scope & Var_perm & Var_GL & Ratio & Gate \\
\midrule
Overall (all layers) & 347.9 & 652.1 & 0.3479 +/- 0.0536 & FAIL \\
Conv2d only & 97.62 & 55.29 & 0.637 & PASS \\
Linear/FC only & 33.84 & 223.52 & 0.133 & FAIL \\
\bottomrule
\end{tabular}
\end{table}

The overall permutation-orbit variance fraction is 0.3479 +/- 0.0536, which is 0.252 below the 0.60 gate threshold. The distribution of per-model ratios (Figure 4) has mean = 0.3479 and std = 0.0536 — consistent across models and not driven by outliers.

\textbf{Layer-type stratification.} The bimodal split between Conv2d (ratio = 0.637) and Linear/FC (ratio = 0.133) is the primary finding from this experiment. The 4.8x difference between these ratios places the two layer types on opposite sides of the 0.60 gate threshold. For Linear (FC) layers, GL-orbit variance is 223.52/33.84 = 6.6x larger than permutation orbit variance.

\textbf{Training trajectory.} The permutation variance ratio decreases monotonically during training: from approximately 0.49 at epoch 0 to approximately 0.28 at epoch 50 (Figure 2). This pattern was not anticipated in the pre-specified hypothesis and was observed consistently across the 1,000 models examined.

\textbf{Correlation with accuracy.} No strong correlation between ratio and final model accuracy was observed (Figure 5, r^2 < 0.05).

\textbf{Gate result: FAIL.} The pre-specified pivot to hybrid encoding was activated.

\subsubsection{5.4 Sub-hypothesis Status Summary}

\begin{table}[htbp]
\centering
\begin{tabular}{llllll}
\toprule
Sub-hypothesis & Gate Type & Result & Key Metric &  &  \\
\midrule
H-E1: Exact symmetry & MUST_WORK & PASS &  & Delta acc & = 0.000000; success rate = 1.0 \\
H-M1: Computability & MUST_WORK & PASS & overhead = 1.167x; unified codebase &  &  \\
H-M2: Variance dominance & MUST_WORK & FAIL & ratio = 0.3479; Conv2d = 0.637, Linear = 0.133 &  &  \\
H-M3: Cross-arch training & — & BLOCKED & Blocked by H-M2 gate failure &  &  \\
H-C1: OVR measurement & — & BLOCKED & Cascaded block &  &  \\
\bottomrule
\end{tabular}
\end{table}

