
\subsubsection{Research Question}

We test whether HumanEval and MBPP produce distinctive evaluation signatures, measured by: (1) ranking divergence (Spearman ρ < 0.8 between benchmark pairs) and (2) distributional divergence (KL divergence > 0.1).

\subsubsection{Data Collection}

\textbf{Feature Extraction:} We extracted 9 execution trace features per model-benchmark pair:
\begin{itemize}
\item \textbf{Correctness:} pass@1, pass@10, pass@100
\item \textbf{Efficiency:} runtime quartiles (25th, 50th, 75th percentile) for passing solutions
\item \textbf{Failure modes:} error counts (syntax, runtime, timeout)
\end{itemize}

\textbf{Data Sources:}
\begin{itemize}
\item Pass@k scores: Published peer-reviewed papers (Chen et al., 2021; Rozière et al., 2023)
\item Runtime measurements: Direct execution on benchmark test suites (Python 3.9, 5-second timeout)
\item Error categorization: Automated classification of failed executions
\end{itemize}

\textbf{Models Evaluated:} 8 code generation models (GPT-4, GPT-3.5-Turbo, CodeLlama-34B, CodeLlama-13B, CodeLlama-7B, StarCoder-15B, DeepSeek-Coder-33B, DeepSeek-Coder-6.7B)

\textbf{Coverage:} 14 model-benchmark pairs total. 6 models have complete data on both HumanEval and MBPP.

\subsubsection{Analysis Methods}

\textbf{Ranking Correlation:} Spearman rank correlation between HumanEval and MBPP model rankings using pass@1 scores. Spearman ρ measures ordinal agreement without requiring identical score values.

\textbf{Distributional Divergence:} Kullback-Leibler divergence between HumanEval and MBPP feature distributions. KL divergence measures statistical dissimilarity independent of ranking agreement.

\textbf{Statistical Significance:} Permutation tests with 10,000 permutations.

\textbf{Success Criteria:} Ranking divergence (ρ < 0.8) AND distributional divergence (KL > 0.1) for at least one benchmark pair.

\subsubsection{Methodological Limitations}

\textbf{Sample Size:} 6 models with HumanEval-MBPP overlap versus originally planned 20+. This limits statistical power for detecting weak ranking divergence.

\textbf{Benchmark Coverage:} Analysis restricted to HumanEval and MBPP. APPS data was unavailable due to dataset API changes during data collection.

\textbf{Metric Focus:} Rankings computed from pass@1 alone. Other metrics (pass@10, pass@100, runtime) were not used for ranking analysis.
