
Dataset documentation frameworks such as the Data Transparency Scorecard (DTS) designate preprocessing and intended-use sections as highest priority, yet no automated pipeline has measured compliance with these standards across multiple ML dataset repositories. We describe an automated DTS-weighted documentation completeness scoring system that queries structured metadata APIs from HuggingFace Hub, OpenML, and UCI ML Repository, mapping retrieved fields to the DTS six-section rubric with inverse-frequency weights. Applied to 758 datasets, the pipeline yields a coverage rate of 91.8\% (696 of 758 datasets scoreable), with a proxy correlation of r = 0.989 (p < 0.001) between weighted and unweighted scores computed on a 120-dataset subsample. However, the two highest-weighted DTS sections -- Preprocessing (weight 1.8) and Uses (weight 1.5) -- exhibit near-zero coverage across all repositories (0.002 and 0.000, respectively), because these sections are not represented as structured API fields. This result reflects a structural gap in repository API design rather than an absence of documentation effort by dataset creators. The pipeline is limited by the use of proxy validation (weighted vs. unweighted correlation) rather than human expert annotation, and by complete scoring failure on the UCI repository due to field naming mismatches. We release the pipeline as an open-source tool for population-scale documentation audits and identify the specific infrastructure changes needed to close the gap between documentation frameworks and repository APIs.
