
We presented an automated DTS-weighted documentation completeness scoring pipeline operating across HuggingFace Hub and OpenML, achieving 91.8\% corpus coverage across 758 datasets with internally consistent scoring (proxy r = 0.989). The pipeline demonstrates that automated cross-repository documentation scoring from structured APIs is technically feasible.

The per-section results reveal a documentation API gap: the DTS sections assigned highest importance weights -- Preprocessing (0.002 coverage) and Uses (0.000 coverage) -- are near-zero across all repositories, not because documentation is absent but because these sections exist as free-text prose rather than structured API fields. Lower-weighted sections such as Motivation (0.547) and Composition (0.267) show substantially higher coverage because their associated information is represented in structured API endpoints.

Several questions remain open. The proxy validation used here does not establish construct validity against human expert judgment; a planned human annotation study (n = 120) is the necessary next step. The UCI repository was not successfully scored due to a field naming mismatch that is engineering-correctable. Three causal hypotheses (H-M1, H-M2, H-M3) remain untested and would require the validated scoring instrument this paper provides.

The pipeline provides measurement infrastructure for future work: fixing the UCI field mapping, conducting human validation, and testing whether documentation completeness predicts dataset adoption. LLM-based parsing of dataset card prose offers a complementary approach to recovering the Preprocessing and Uses signals that structured APIs currently do not expose.
