
\subsubsection{6.1 Interpretation of Findings}

The primary finding is a negative result: energy redistribution toward slow eigenmodes does not occur under projection-only LoRA. This eliminates the Eigenmode Utilization Hypothesis as an explanation for potential beyond-horizon task success.

The per-layer analysis reveals that Mamba-1.4B allocates only 0.002\% of total state energy to slow modes, concentrated in layers 18 and 19. The architecture favors fast-decaying eigenmodes rather than long-term memory.

These findings provide evidence that projection-only LoRA is bounded by the intrinsic spectral horizon H\_spec. However, direct verification of task failure at L > H\_spec was not performed because H-M4 was not executed.

\subsubsection{6.2 Limitations}

\textbf{WikiText-103 as Memory Proxy:} Language modeling perplexity was used rather than a controlled synthetic task with explicit dependency lengths. The perplexity degradation pattern provides indirect evidence but does not constitute direct verification of task failure at the spectral boundary.

\textbf{Single Model Family:} Experiments focused on Mamba-1.4B and Mamba-370M. Generalization to other SSM architectures (RWKV, RetNet, Mamba-2) was not empirically validated.

\textbf{H-M4 Not Executed:} The discriminative MQAR test that would directly verify task failure when L > H\_spec under projection-only LoRA was not completed. The H-M3 negative result partially addresses the hypothesis discrimination but does not provide direct task-failure evidence.

\textbf{Proof-of-Concept Training Configuration:} A single epoch with limited training sequences was used. The configuration was sufficient for mechanism validation (eigenvalue preservation and energy measurement) but does not represent full-scale training.

\textbf{SSM-Core LoRA Not Tested:} LoRA targeting A\_log or Delta parameters was not implemented. Prediction P3 (that SSM-core LoRA can extend H\_spec) remains untested.

\subsubsection{6.3 Connection to Prior Work}

The findings provide a potential explanation for observations in prior work:

\begin{itemize}
\item SSM-PEFT reports that LoRA is ineffective on SSM modules. The present results suggest this is because eigenvalues are frozen and energy redistribution does not occur.
\item MambaPEFT reports task-dependent LoRA success. This may correspond to tasks within versus beyond H\_spec, though this interpretation requires validation.
\end{itemize}

\subsubsection{6.4 Scope Conditions}

Results hold for:
\begin{itemize}
\item Mamba-1.4B architecture with projection-only LoRA targeting in\_proj and x\_proj
\item Language modeling evaluation on WikiText-103
\item The specific pretrained checkpoint (state-spaces/mamba-1.4b-hf)
\end{itemize}

Results may not generalize to:
\begin{itemize}
\item Other SSM architectures (RWKV, Mamba-2, hybrid Transformer-SSM models)
\item SSM-core LoRA or full fine-tuning
\item Tasks other than language modeling
\item Different pretrained checkpoints
\end{itemize}
