
\subsubsection{5.1 H-E1: Spectral Horizon Stability}

\textbf{Research Question:} Is H\_spec a stable, input-independent property?

\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{llll}
\toprule
Metric & Value & Threshold & Result \\
\midrule
CV(H\_spec) & 2.22 x $10^-16$ & < 0.3 & PASS \\
Mean H\_spec & 256.18 tokens & - & - \\
Std H\_spec & 5.68 x $10^-14$ & - & - \\
Valid samples & 1000/1000 & - & - \\
\bottomrule
\end{tabular}
}
\end{table}

The coefficient of variation is effectively zero, indicating that H\_spec is determined solely by the pretrained A matrix weights and is independent of input sequences.

\textbf{Cross-Scale Observation:} Mamba-370M yielded H\_spec = 162,605 tokens, larger than Mamba-1.4B (256.18 tokens). This non-monotonic relationship was unexpected and suggests H\_spec is determined by pretraining dynamics rather than model capacity alone.

\subsubsection{5.2 H-M1: Eigenvalue-Based Memory Prediction}

\textbf{Research Question:} Does H\_spec predict actual memory behavior on real text?

\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{ll}
\toprule
Context Length & Perplexity \\
\midrule
25 tokens & 83.26 \\
64 tokens & 36.40 \\
128 tokens & 23.75 \\
256 tokens & 17.89 \\
512 tokens & 14.41 \\
1024 tokens & 12.22 \\
\bottomrule
\end{tabular}
}
\end{table}

\textbf{Degradation Ratio:} 3.03 (threshold: > 1.1) --- PASS

Perplexity improved rapidly as context approached H\_spec (256 tokens), then showed diminishing returns beyond H\_spec. The improvement from 25 to 256 tokens was 4.65x, while the improvement from 256 to 1024 tokens was 1.47x. This pattern indicates that H\_spec predicts the point at which memory becomes the limiting factor for language modeling performance.

\subsubsection{5.3 H-M2: Eigenvalue Preservation Under LoRA}

\textbf{Research Question:} Does projection-only LoRA preserve SSM eigenvalues?

\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{llll}
\toprule
Metric & Pre-Training & Post-Training & Change \\
\midrule
H\_spec & 256.43 tokens & 256.43 tokens & 0.0000\% \\
A\_log max diff & - & 0.0 & Frozen \\
Eigenvalue correlation & - & 1.0000 & Perfect \\
Validation perplexity & - & 15.58 & - \\
\bottomrule
\end{tabular}
}
\end{table}

\textbf{Result:} |delta H\_spec| = 0.0\% (threshold: < 10\%) --- PASS

The A\_log parameters remained completely frozen during training. Eigenvalue preservation was exact, not approximate. This confirms architectural isolation between projection parameters and SSM core parameters.

\subsubsection{5.4 H-M3: Eigenmode Energy Redistribution}

\textbf{Research Question:} Can projection-only LoRA redistribute energy to slow eigenmodes?

\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{llll}
\toprule
Metric & Pre-Training & Post-Training & Change \\
\midrule
Slow mode fraction & 1.97 x $10^-5$ & 1.91 x $10^-5$ & -5.93 x $10^-7$ \\
delta E (nats) & - & 5.93 x $10^-7$ & 6 orders below threshold \\
Perplexity & - & 14.35 & - \\
\bottomrule
\end{tabular}
}
\end{table}

\textbf{Result:} delta E = 5.93 x $10^-7$ nats (threshold: > 0.1 nats) --- FAIL

Energy redistribution was negligible. Per-layer analysis revealed:

\begin{itemize}
\item Only 2 of 48 layers (layers 18 and 19) contain slow eigenmodes (|lambda| > 0.99)
\item Layer 18: 0.044\% slow mode fraction
\item Layer 19: 0.050\% slow mode fraction
\item All other 46 layers: 0.0\% slow mode fraction
\item Total slow mode capacity: 0.00197\% of state energy
\end{itemize}

The pretrained Mamba-1.4B architecture allocates minimal energy to slow-decaying modes. There is insufficient slow-mode capacity for meaningful redistribution.

\subsubsection{5.5 Summary of Results}

\begin{table*}[htbp]
\centering
\resizebox{\dimexpr 2\columnwidth+\columnsep\relax}{!}{%
\begin{tabular}{llllllll}
\toprule
Hypothesis & Gate & Metric & Threshold & Actual & Result &  &  \\
\midrule
H-E1 & MUST\_WORK & CV(H\_spec) & < 0.3 & 2.22 x $10^-16$ & PASS &  &  \\
H-M1 & MUST\_WORK & Degradation ratio & > 1.1 & 3.03 & PASS &  &  \\
H-M2 & MUST\_WORK &  & delta H\_spec &  & < 10\% & 0.0\% & PASS \\
H-M3 & SHOULD\_WORK & delta E & > 0.1 & 5.93 x $10^-7$ & FAIL &  &  \\
\bottomrule
\end{tabular}
}
\end{table*}

The three MUST\_WORK hypotheses passed. The SHOULD\_WORK hypothesis (H-M3) failed, which eliminates the EUH mechanism and provides evidence for MHSH.
