
\subsubsection{Overview}

CAWE consists of three components: (1) architecture-specific tokenizers that project weights to D=128 dimensional token sequences, (2) a shared NFT backbone with permutation-equivariant attention, and (3) a regression head for generalization gap prediction.

\subsubsection{Architecture-Specific Tokenizers}

\textbf{CNN Tokenizer}: For each convolutional layer with weights $W \in \mathbb{R}^{C_\text{out} \times C_\text{in} \times K_h \times K_w}$, the kernel is flattened and projected to dimension 128 via a learned linear layer. Tokens from all layers are concatenated to form a sequence of length $T_\text{CNN}$.

\textbf{Transformer Tokenizer}: For each Vision Transformer layer, Query, Key, and Value weight matrices are extracted from attention heads, flattened, and projected to dimension 128. Tokens from all layers and heads are concatenated to form a sequence of length $T_\text{ViT}$.

\textbf{MLP Tokenizer}: For each fully-connected layer with weights $W \in \mathbb{R}^{n_\text{out} \times n_\text{in}}$, the weight matrix is flattened and projected to dimension 128. Tokens from all layers are concatenated to form a sequence of length $T_\text{MLP}$.

The shared token dimension D=128 was selected based on preliminary observations that D=64 showed insufficient capacity while D=256 provided no performance improvement over D=128 in pilot experiments.

\subsubsection{NFT Backbone}

The tokenized sequences are processed by a Neural Functional Transformer using the nfn library implementation \citep{zhou2023meta}. The configuration uses 4 attention layers, 8 attention heads, hidden dimension 256, and dropout 0.1. NFT attention provides permutation-equivariance, meaning that reordering neurons in the input network (which does not change network function) produces equivalent representations.

\subsubsection{Regression Head}

The NFT produces an embedding which is passed through a two-layer MLP with ReLU activation to predict a scalar generalization gap (test accuracy - train accuracy). The model is trained with mean squared error loss.

\subsubsection{Training Configuration}

Optimizer: AdamW with learning rate 1e-4 and weight decay 1e-2. Batch size: 16. Maximum epochs: 100 with early stopping (patience 10 epochs on validation Spearman $\rho$). Data split: 150 training, 30 validation, 30 test models, stratified by architecture family. Random seed: 42.
