
\subsubsection{5.1 H-E1: Extraction Feasibility}

Linguistic markers were successfully extracted with the following distributional properties:

\begin{table*}[htbp]
\centering
\resizebox{\dimexpr 2\columnwidth+\columnsep\relax}{!}{%
\begin{tabular}{lllll}
\toprule
Marker Type & Mean & SD & CV & Precision \\
\midrule
Modal verbs & 2.911 & 2.272 & 0.781 & 100\% \\
Hedging & 0.542 & 0.416 & 0.768 & 100\% \\
Alternatives & 0.089 & 0.112 & 1.258 & 100\% \\
\bottomrule
\end{tabular}
}
\end{table*}

Modal verb CV (0.781) exceeded the threshold (0.3) by 161\%. Extraction precision was 100\% with 98.5\% recall. Cross-split consistency was high (train CV=0.783, test CV=0.781, difference=0.002). H-E1 criteria were satisfied.

\subsubsection{5.2 H-M: Construct Validation}

\textbf{Effect Size.} Paired t-test results on 169,352 pairs:

\begin{table*}[htbp]
\centering
\resizebox{\dimexpr 2\columnwidth+\columnsep\relax}{!}{%
\begin{tabular}{llllll}
\toprule
Marker & Chosen Mean & Rejected Mean & Difference & Cohen's d & p-value \\
\midrule
Modal verbs & 2.894 & 2.928 & -0.034 & -0.0181 & <0.001 \\
Hedging & 0.538 & 0.546 & -0.008 & -0.0192 & <0.001 \\
Alternatives & 0.087 & 0.091 & -0.004 & -0.0356 & <0.001 \\
\bottomrule
\end{tabular}
}
\end{table*}

All three markers achieved statistical significance (p<0.001) but effect sizes were 76-88\% below the threshold (d=0.15). For modal verbs, the observed d=-0.0181 represents 12\% of the required threshold. The absolute frequency difference was 0.034 per 100 words (1.2\% difference). Direction was consistent with theoretical predictions (chosen < rejected).

\textbf{Internal Consistency.} Cronbach's $\alpha$ across the three marker types was 0.42, falling 40\% below the acceptable threshold (0.7). Mean inter-item correlation was 0.29, with pairwise correlations ranging from 0.22 to 0.35.

\textbf{Cross-Split Replication.} Separate analyses by data split:

\begin{table*}[htbp]
\centering
\resizebox{\dimexpr 2\columnwidth+\columnsep\relax}{!}{%
\begin{tabular}{lllll}
\toprule
Split & N & Cohen's d & p-value & Passed \\
\midrule
Train & 160,800 & -0.0187 & <0.001 & No \\
Test & 8,552 & -0.0067 & 0.536 & No \\
\bottomrule
\end{tabular}
}
\end{table*}

Neither split achieved the effect size threshold. The test split did not reach statistical significance despite 95\% power to detect d=0.15 effects.

\textbf{Summary.} All four H-M criteria failed:
\begin{itemize}
\item Effect size: d=-0.018 vs. required $d\geq 0.15$ (88\% below threshold)
\item Internal consistency: $\alpha =0.42$ vs. required $\alpha$>0.7 (40\% below threshold)
\item Replication: 0/2 splits passed
\item Practical significance: 1.2\% frequency difference
\end{itemize}
