
Three sub-hypotheses are evaluated:

\begin{itemize}
\item \textbf{H-E1 (existence):} Do stylistic preference coefficients show nominal significance and feature orthogonality in HH-RLHF? (MUST_WORK gate)
\item \textbf{H-M1 (mechanism, geometric):} Is the annotation preference gradient aligned with the AI-typicality direction in embedding space? (MUST_WORK gate)
\item \textbf{H-M2 (mechanism, coefficient):} Do stylistic preference coefficients shift directionally and significantly across annotation strata? (SHOULD_WORK gate)
\end{itemize}

Two additional sub-hypotheses (H-M3: reward model behavioral divergence; H-M4: benchmark degradation) were fully specified but not executed within this project.

\subsubsection{4.1 Dataset Summary}

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
Dataset & Rows & Strata / Groups & Design \\
\midrule
HH-RLHF & 160,800 & 3 strata × 53,600 rows (index-based) & Round-stratified logistic regression \\
WebGPT & 19,578 & Between-group tercile & Between-group projection regression \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{4.2 Baselines and Controls}

\begin{itemize}
\item \textbf{Random temporal split:} Rules out sampling artifacts.
\item \textbf{Q_early predictor (frozen):} Enables stylistic-quality decomposition.
\item \textbf{Placebo AI-typicality vector:} 200-iteration permutation test for discriminant validity.
\end{itemize}

\subsubsection{4.3 Evaluation Criteria}

\begin{table}[htbp]
\centering
\begin{tabular}{lll}
\toprule
Metric & Pre-registered Threshold & Sub-hypothesis \\
\midrule
n_directional & ≥ 2 of 3 non-overlapping CIs & H-M2 \\
sign_consistent & all Δβ > 0 & H-M2 (diagnostic) \\
β_exposure effect size & ≥ 0.1 SD per 1,000 tokens & H-M1 \\
Placebo empirical p & > 0.05 & H-M1 (discriminant validity) \\
β_L Bonferroni p & < 0.0167 & H-E1 \\
Round × ambiguity interaction p & < 0.0167 & H-E1 \\
\bottomrule
\end{tabular}
\end{table}

Multiple comparisons across three hypothesis tests are addressed with a Bonferroni-corrected family-wise threshold of α = 0.05/3 ≈ 0.017.

\subsubsection{4.4 Implementation}

scikit-learn LogisticRegression (C=1.0, solver=lbfgs, max_iter=1000, class_weight=balanced, random_state=42); 75/25 train/test split; 2,000 stratified bootstrap resamples; sentence-transformers all-MiniLM-L6-v2 (frozen); single NVIDIA H100 NVL GPU (95,830 MiB).

