
\subsubsection{Research Questions}

The experiments address three research questions corresponding to the three executed sub-hypotheses (H-E1, H-M1, H-M2). Sub-hypotheses H-M3 ($margin\times method$ interaction in mixed-effects logistic regression) and H-M4 (cosine projection of logit deltas onto the dominant perturbation eigenvector) were not executed; their planned analyses remain inconclusive.

\subsubsection{Evaluation Metrics}

For RQ1 (H-E1): $\beta _1$ sign and magnitude, Wald p-value ($\alpha$ = 0.005), AUROC per benchmark (threshold 0.75), partial $\eta^2$ (threshold 0.06). Gate type: MUST\_WORK.

For RQ2 (H-M1): anisotropy ratio vs. isotropic control ($\approx 1.13)$, one-tailed t-test significance ($\alpha$ = 0.05), minimum 2 of 3 model families passing both criteria. Gate type: MUST\_WORK.

For RQ3 (H-M2): per-quintile logit delta variances, one-tailed Welch's t-test for DPO Q1 > SFT Q1, Cohen's d. Gate type: SHOULD\_WORK (null result acceptable; pipeline continues).

\subsubsection{Implementation Details}

For each item, 4D log-probability vectors over answer tokens {A, B, C, D} are extracted from both base and aligned models. Logit extraction uses cached .npy files (18 files for H-E1: 6 per completed pair $\times$ 3 datasets). The anisotropy analysis reuses the H-E1 logprob cache without re-running model inference. The quintile variance analysis uses the same cache. All code is implemented in Python with numpy, scipy, and sklearn; 45 unit tests pass for H-M1 and 23 for H-M2.

