
\subsubsection{4.1 Experiment 1: Capability Invariance Test (h-e1)}

\textbf{Objective:} Validate assumption A1 (policy-layer modulation preserves capability).

\textbf{Protocol:}
\begin{itemize}
\item Model: Claude 3 Opus, policy-layer modulation via system prompts, temperature 0.0
\item Compliance Levels: λ ∈ {0.2, 0.4, 0.6, 0.8, 1.0}
\item Datasets: MMLU (57 subjects, 14,042 test questions, 4-shot prompting), HumanEval (164 problems, pass@1 metric)
\item Measurements: Accuracy per λ condition
\item Statistical Tests: ICC₂ > 0.95, ANOVA p > 0.05, Cohen's f < 0.10
\item Gate Logic: All three criteria must pass
\end{itemize}

\textbf{Expected Cost:} ~$1,620 for full evaluation. \textbf{Expected Runtime:} ~4 hours.

\textbf{Status:} Infrastructure code validated and datasets verified. Not executed due to API resource constraint (ANTHROPIC_API_KEY unavailable).

\subsubsection{4.2 Experiment 2: Coupling Characterization (h-m1)}

\textbf{Objective:} Test whether ACE-HOR coupling exists and which model (M1/M2/M3) fits.

\textbf{Prerequisites:} h-e1 must pass.

\textbf{Protocol:}
\begin{itemize}
\item Task: Human participants review AI-generated outputs across λ conditions
\item Seeded Errors: 20\% within-distribution errors
\item HOR Measurement: d' = Z(hit_rate) - Z(false_alarm_rate) per λ condition
\item Model Fitting: M1 (Linear), M2 (Quadratic), M3 (Bifurcation)
\item Model Selection: AIC comparison, require R² ≥ 0.6
\end{itemize}

\textbf{Status:} Not started (prerequisite h-e1 not completed).

\subsubsection{4.3 Experiment 3: Mediation Test (h-m2, h-m3)}

\textbf{Objective:} Test whether HOR degradation is mediated by perceived reliability.

\textbf{Prerequisites:} h-e1 and h-m1 must pass.

\textbf{Design:} 2×2 randomized experiment with factors: ACE (λ = 0.4 vs. 0.8) and Trust Display (Standard vs. Randomized).

\textbf{Hypothesis:} If automation bias → HOR_randomized > HOR_standard. Threshold: HOR difference ≥ 10\%, p < 0.05.

\textbf{Status:} Not started (prerequisites not completed).

\subsubsection{4.4 Experiment 4: Metacognitive Intervention (h-m4)}

\textbf{Objective:} Test whether designed disagreement sustains HOR via increased learning rate k₂.

\textbf{Prerequisites:} h-e1, h-m1, h-m2, h-m3 must pass.

\textbf{Design:} Intervention vs. control, longitudinal tracking at t ∈ {1 hour, 1 day, 1 week, 1 month}.

\textbf{Success Criterion:} k₂_intervention / k₂_control ≥ 1.30 with 95\% Bayesian credible interval excluding 1.0.

\textbf{Status:} Not started (prerequisites not completed).
