
Three research questions guided the experiments:

\textbf{RQ1 (Registry Feasibility):} Can a registry of at least 200 LLMs with at least 3 non-zero-variance curation documentation features be assembled from publicly available data?

\textbf{RQ2 (Scale Baseline):} How much V2 benchmark variance is explained by scale and architecture alone?

\textbf{RQ3 (Documentation Effect):} After scale controls, does doc\_score predict benchmark performance, and in what direction?

\subsubsection{4.1 Dataset}

\begin{table}[htbp]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tabular}{ll}
\toprule
Stage & Count \\
\midrule
Leaderboard rows (raw) & 4,497 \\
After benchmark coverage filter (>= 4/6) & 4,493 \\
Targeted family models prioritized & 114 \\
Model cards retrieved (total) & 3,749 \\
Models with doc\_score >= 1 & 158 \\
\bottomrule
\end{tabular}
}
\end{table}

The 114 models matching targeted family prefixes were prioritized for retrieval; remaining cards were retrieved within the same API budget window until the rate limit was reached.

V2 benchmarks (IFEval, BBH, MATH Level 5, GPQA, MUSR, MMLU-PRO) were used rather than V1 because they measure harder capabilities less susceptible to scale saturation.

\subsubsection{4.2 Baselines}

\textbf{Scale Baseline:} OLS with log(params) + log(tokens) + architecture family fixed effects, establishing how much variance is attributable to scale.

\textbf{Scale + Documentation (Proposed):} Adds doc\_score to the scale baseline.

\subsubsection{4.3 Evaluation Metrics}

Primary metrics: n\_analyzable, n\_features\_with\_variance, R-squared\_baseline, beta\_docs, p\_value\_docs, delta\_R-squared. Statistical significance threshold: alpha = 0.05 with HC3 robust standard errors. Nested model comparison: the F-test for a single added regressor is equivalent to the squared t-statistic for beta\_docs.

\subsubsection{4.4 Implementation Details}

Python 3.10, pandas, statsmodels 0.14, huggingface\_hub 0.20. All regex patterns were pre-registered before data access. Total H-E1-v2 runtime was approximately 6.7 hours, dominated by API rate-limited card retrieval. The analysis was CPU-only (no GPU required for OLS).
