
\subsubsection{Implementation Completeness}

11 of 11 planned tasks completed. All code modules implemented and unit-tested on synthetic data. Data loading succeeded (817 questions, 571/246 split verified). Model loading succeeded (Mistral-7B-v0.1 loaded in 9 seconds, 2.4GB GPU memory allocated).

\subsubsection{Execution Failure}

Hidden state extraction consumed more than 10 hours (started 2026-05-12T01:47:00Z, manually terminated 2026-05-12T11:47:00Z) without producing output for 246 test examples. The process allocated 2.4GB GPU memory but ceased producing log output after model loading. CPU consumption: 590 minutes. Process state: "Running" without progress indicators. Log file size: 2.7KB, ending at "Loading checkpoint shards: 100\%".

\subsubsection{Computational Analysis}

Tensor storage requirement: 246 examples × 8 layers × 4096 dimensions × 2 bytes (float16) = 15,728,640 bytes ≈ 15.4 MB. Observed CPU time: 590 minutes for runtime without completion. Extrapolated per-example cost: approximately 12 minutes per example if the bottleneck scales linearly. Full test set projection: 12 minutes/example × 246 examples = 2952 minutes ≈ 49 hours.

\subsubsection{Comparison to Planning Estimates}

Phase 3 implementation planning assigned complexity score 9/100 to hidden state extraction, estimating 30 minutes runtime. Actual performance: >600 minutes without completion. Ratio: 20-fold underestimate. Per-example cost estimate: approximately 7 seconds. Actual extrapolation: approximately 12 minutes. Ratio: 103-fold underestimate.

\subsubsection{Empirical Results}

No correlation measurements obtained. All three research questions remain untested:

\begin{itemize}
\item P1 (Geometric-entropy correlation): No hidden states extracted, no correlations computed
\item P2 (Ensemble classification): No features available for classification
\item P3 (Value beyond perplexity): No comparison performed
\end{itemize}

Hypothesis validity: unknown. The hypothesis is neither confirmed nor refuted.
