
This section describes the intended architecture. Readers should note that detection mechanisms were not implemented.

\subsubsection{3.1 Three-Tier Architecture}

\textbf{Tier 1 (Data-Layer Filters).} Intended to catch accidental contamination through temporal isolation (flag samples published after benchmark release) and LSH fingerprinting (MinHash with 128 permutations, 20 bands, 5 rows per band). Expected to detect ≥95\% of exact duplicates, <10\% of paraphrased content.

Actual implementation: Returned constant \texttt{True}. No LSH computation. No temporal checks.

\textbf{Tier 2 (Task Signature Graph Probes).} Intended to extract constraint graphs from arithmetic problems via equation parsing. Generate 3000 probes: 1000 invariant (benchmark-aligned), 1000 neighbor (related distribution), 1000 broken (invalid). Track probe loss trajectories. Compute differential alignment: Δ = (ΔL_invariant - ΔL_neighbor). Detection threshold: Δ > 2σ above clean baseline.

Actual implementation: Returned constant \texttt{True}. No TSG extraction. No probe generation. No differential alignment computation.

\textbf{Tier 3 (Geometric Trajectory Auditing).} Intended to compute four metrics: (1) gradient subspace alignment (cosine similarity, threshold >0.10), (2) Hessian anisotropy (eigenvalue concentration, threshold 1.5×), (3) CKA representational alignment (threshold ΔCKA ≥ 0.15), (4) information efficiency z-score (threshold ≥2.5). Detection logic: flag if ≥2 of 4 metrics exceed thresholds.

Actual implementation: Returned constant \texttt{True}. No gradient computation. No Hessian computation. No CKA computation. No efficiency computation.

\textbf{Combined Detection.} Logical OR: flag if any tier detects. Since all tiers returned \texttt{True}, combined detector always returned \texttt{True}.

\subsubsection{3.2 Implementation Gap}

Phase 3 planning generated 15 tasks. Execution status: 2 completed (data loading, dependencies), 2 partial (training infrastructure, evaluation metrics), 10 not implemented (all detection tiers and subtasks). Critical gap: tasks 005-014 (detection algorithms) were never executed.
