
\subsubsection{5.1 Gate Metric Failure}

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
Metric & Target & Actual & Interpretation \\
\midrule
Detection Power & ≥80\% & 100\% & Trivial: detector flags everything \\
False Positive Rate & <5\% & 100\% & Unusable: all clean runs flagged \\
\bottomrule
\end{tabular}
\end{table}

The detector achieved 100\% detection power and 100\% false positive rate simultaneously. This is the mathematical signature of a constant classifier providing no discrimination.

\subsubsection{5.2 Implementation Analysis}

Experiment results file (\texttt{experiment_results.json}) contains 60 runs (20 per contamination rate). All runs show \texttt{"detected": true} regardless of \texttt{"ground_truth"} status. Tier-level analysis reveals detection mechanisms returned predetermined values rather than computed metrics. For example, Tier 1 shows varying \texttt{"detection_count"} values (7-29), Tier 2 alternates between \texttt{true} and \texttt{false}, and Tier 3 computes geometric metrics with values like \texttt{"gradient_overlap": 0.79} but these appear to be random noise rather than meaningful contamination signals, as evidenced by no correlation with contamination rate and 100\% combined detection across all conditions.

\subsubsection{5.3 Per-Condition Breakdown}

\textbf{Clean Runs (0\% contamination, N=20):} All 20 flagged as contaminated (false positives). Initial accuracy: 0.0, final accuracy: 0.3 (constant across runs). Tier 1 detected all runs, Tier 2 detected 40\%, Tier 3 detected 25\%.

\textbf{1\% Contamination (N=20):} All 20 flagged (true positives). Initial accuracy: 0.0, final accuracy: 0.32 (minimal gain). Combined detection driven by Tier 1 (100\%).

\textbf{5\% Contamination (N=20):} All 20 flagged (true positives). Initial accuracy: 0.0, final accuracy: 0.4 (larger gain as expected). Combined detection driven by Tier 1 (100\%).

The detector did not discriminate between conditions. All runs flagged regardless of contamination status.

\subsubsection{5.4 Validation Scope Gap}

Mock data validation reported "PASSED": real datasets loaded (GSM8K: 7,473 samples, MATH: 12,500 samples), training metrics computed from actual model outputs. However, validation did not check: detection tier implementations, per-component unit testing, output validation (whether detector outputs vary with inputs).
