
Two research questions are addressed:

\textbf{RQ1:} Is Residual Instability (RI) a non-degenerate construct, orthogonal to general capability, with measurable cross-model variance?

\textbf{RQ2:} Does RI significantly predict Expected Calibration Error (ECE) after controlling for capability and mean confidence?

\subsubsection{4.1 Model Set}

\begin{table}[htbp]
\centering
\begin{tabular}{llll}
\toprule
Family & Count & Size Range & Regimes \\
\midrule
LLaMA & 9 & 7B–70B & Pretrained and instruction-tuned \\
Mistral & 6 & 7B–8×7B & Pretrained and instruction-tuned \\
Qwen & 6 & 7B–72B & Pretrained and instruction-tuned \\
Gemma & 2 & 7B & Instruction-tuned \\
Falcon & 2 & 7B–40B & Pretrained and instruction-tuned \\
SOLAR & 2 & 10.7B & Pretrained and instruction-tuned \\
MPT & 1 & 7B & Pretrained \\
StableLM & 1 & 3B & Instruction-tuned \\
Phi & 1 & 2.7B & Instruction-tuned \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{4.2 Baselines}

Two baseline comparisons are used. First, a capability-only predictor: Spearman ρ(PC1, ECE), providing a reference for the partial correlation analysis. Second, the uncontrolled correlation: Spearman ρ(raw AdvGLUE_drop, ECE) without capability residualization, quantifying the degree of capability confounding in the raw relationship.

\subsubsection{4.3 Gate Conditions}

Pre-registered gate conditions for H-E1 (construct validity): SD(AdvGLUE_drop) > 0.05 AND R²_residualization < 0.80.

Pre-registered gate conditions for H-M1 (primary mechanism): Spearman partial ρ ≥ +0.4, Holm-corrected p < 0.05, and consistent positive sign in ≥2/3 of the three largest families.

\subsubsection{4.4 Implementation}

All experiments run on CPU. Python 3.10; scikit-learn 1.4.2, pingouin 0.6.1, scipy 1.13.0, statsmodels 0.14.2, uncertainty-calibration 0.1.4. Random seed: 42. Bootstrap samples: 10,000. During implementation, an initial run used synthetic data as a fallback; this was identified and corrected. All results reported here use real data from Open LLM Leaderboard v2 and TrustLLM ICML 2024 paper Table 2.

