
\subsubsection{6.1 The Zero Type Stratum}

The finding that CodeLlama-7B produces zero type errors across 2,680 completions from 134 HumanEval problems defines a scope boundary for type-checking repair applied to this model-benchmark pair.

Two explanations are plausible. First, CodeLlama-7B is trained on GitHub Python repositories where annotation-free, dynamically-typed code predominates, so the model's output distribution reflects its training distribution and generates annotation-free Python by default. Second, the FMD priority chain classifies failures as syntax first; with 97.5\% of completions failing at the syntax level, type errors in post-syntax-correction code would be masked by the syntax classifier. Both explanations likely contribute.

These explanations are distinguishable: applying mypy to SynCode-generated pools (which have reduced syntax failure rates) should reveal whether type errors emerge when the syntax layer is partially removed.

This finding is relevant to the iterative feedback repair literature, which has cited mypy and similar static analysis tools as feedback mechanisms for LLM repair. For CodeLlama-7B on HumanEval, this feedback channel is empirically vacuous.

\subsubsection{6.2 SynCode: Directional Effect Without Statistical Significance}

The h-m1 result (delta_ast = 0.075, p = 0.1186) reflects insufficient statistical power at N = 20, not necessarily the absence of an effect. Power analysis indicates approximately 25\% power at this sample size and effect size. The same delta_ast value was observed in two independent experiments (h-e1 and h-m1), which provides some corroborating evidence for the stability of the effect. A 7.5 percentage point reduction in AST failure rate corresponds to approximately 1-2 additional correct completions per 20 samples per problem.

The \texttt{constraint_active = False} observation in h-e1 raises a question about whether the measured improvement is attributable to full CFG enforcement or to partial grammar masking effects. This distinction was not resolved by the current experiments.

\subsubsection{6.3 Z3: Established Scope, Untested Repair}

Z3 eligibility (33\% of HumanEval) is established. The SMT repair mechanism itself was not executed (h-m3 was not started). The complementarity claim for the SynCode-Z3 pair therefore remains inconclusive. All infrastructure required for this experiment (TransitionExtractor, CScoreCalculator, Z3EligibilityChecker) has been implemented and tested.

\subsubsection{6.4 Pipeline Reduction}

The three-method formal repair pipeline (SynCode + mypy + Z3) for CodeLlama-7B on HumanEval is effectively a two-method pipeline, because mypy does not activate. The direct SynCode-Z3 path remains theoretically valid and untested. These two methods address structurally distinct failure channels: syntax invalidity versus arithmetic constraint unsatisfiability.

\subsubsection{6.5 Limitations}

\textbf{L1: h-m3 and h-m4 not executed.} The SynCode-Z3 complementarity experiment (h-m3) and the feature-aware routing experiment (h-m4) were not executed. The core complementarity claim remains untested. All infrastructure is implemented; h-m3 requires approximately one additional hour of GPU time with existing code.

\textbf{L2: SynCode statistical significance pending.} The directional result (delta_ast = 0.075) is not statistically significant at N = 20. Full confirmation requires N >= 60 problems.

\textbf{L3: Single model, single benchmark.} All results are for CodeLlama-7B (7B parameters) on HumanEval. Larger models, instruction-tuned variants, or annotation-prompted models may exhibit different failure distributions. The zero type stratum finding is model-specific and should not be generalized without additional measurement.

\textbf{L4: Z3 eligibility heuristic inconsistency.} h-m2 used a stricter annotation-based heuristic than h-e1 (assertion-scanning). The h-e1 measurement (33\%) is the correct value.

\textbf{L5: constraint_active flag.} SynCode's internal \texttt{constraint_active} flag was False during h-e1 generation, indicating that not all token-level grammar constraints were enforced. The mechanism by which delta_ast = 0.075 was achieved under partial constraint enforcement is not fully characterized.
