
Three experiments address the following questions: (1) Are all three formal repair tools operational in a unified Python-native pipeline? (2) Does SynCode produce measurable FMD improvement? (3) What is the FMD of CodeLlama-7B on HumanEval, and which repair channels can activate?

\subsubsection{4.1 Model and Infrastructure}

\textbf{Base model.} CodeLlama-7B (\texttt{codellama/CodeLlama-7b-hf}), loaded with \texttt{device_map='auto'} and \texttt{torch_dtype=float16}.

\textbf{Software.} Python 3.10, SynCode v0.4.16, z3-solver v4.16.0.0, mypy v1.20.2, evalplus v0.3.1. All tools pip-installed in a single conda environment.

\textbf{Benchmark.} HumanEval (Chen et al., 2021), 164 Python function-level synthesis problems. Experiments h-e1 and h-m1 use a 20-problem subset (problems 0-19 by lexicographic sort). Experiment h-m2 uses a 134-problem baseline pool.

\textbf{Frozen pool.} N = 20 completions per problem, fixed seeds (seed_i = problem_idx * 100 + sample_idx), temperature = 0.8, max_new_tokens = 512, top_p = 0.95.

\subsubsection{4.2 Experiment 1: Unified Pipeline Operationality (h-e1)}

\textbf{Gate type:} MUST_WORK. \textbf{Success criteria:} delta_ast > 0; z3_eligibility_rate >= 0.15; mypy_structured_rate >= 0.90.

\subsubsection{4.3 Experiment 2: SynCode Mechanism Test (h-m1)}

\textbf{Gate type:} MUST_WORK. \textbf{Success criteria:} delta_ast > 0; bootstrap CI lower bound > 0.

\textbf{Statistical power note.} N = 20 problems yields approximately 25\% power at delta = 0.075. N >= 60 is required for 80\% power.

\subsubsection{4.4 Experiment 3: Failure Mode Distribution and Channel Scope (h-m2)}

\textbf{Gate type:} SHOULD_WORK. \textbf{Success criteria:} C_score > 0, bootstrap p < 0.0167; delta_P(Z3_eligible | post-mypy) > 0.05.

\subsubsection{4.5 Metrics}

\begin{table}[htbp]
\centering
\begin{tabular}{lllll}
\toprule
Metric & Formula & Interpretation &  &  \\
\midrule
delta_ast & mean(AST_fail_baseline) - mean(AST_fail_SynCode) & SynCode AST failure reduction &  &  \\
z3_eligibility_rate & Z3-eligible problems / all problems & Z3 applicability scope &  &  \\
mypy_structured_rate & Completions with mypy output / all completions & mypy operational rate &  &  \\
C_score & E[J_indep] - J_obs(F_A, F_B) & Complementarity above independence &  &  \\
delta_P(Z3 eligible) & P(eligible & post-mypy) - P(eligible & baseline) & Cascade expansion rate \\
\bottomrule
\end{tabular}
\end{table}
