
LoRA adapters for large language models (LLMs) are typically trained with full KV cache access but deployed under eviction policies that discard 50–75\% of cached token entries. This training-inference distribution mismatch is not directly addressed by any existing fine-tuning method. This paper proposes Eviction-Aware LoRA, which injects H2O eviction masks into the LoRA forward pass during training, exposing adapter parameters to the evicted-cache attention distribution they will encounter at inference. Experiments on a GPT-2 proxy model (117M parameters, 30 training steps, 200 LongAlpaca-12k samples) show that eviction-aware training produces adapter weights with a mean cosine similarity of 0.053 relative to sequential-baseline adapters (range: −0.578 to +0.469 across 24 layers), with all 24 layers falling below the 0.95 divergence threshold. Attention entropy analysis on 5 synthetic samples shows significant divergence (paired t-test, p < 0.05) in 8 of 12 transformer layers, concentrated in middle layers 4–11; no significant divergence was observed in lower layers 0–3 by either entropy or heavy-hitter concentration metrics. These mechanistic results confirm that eviction-aware training alters the gradient signal reaching LoRA parameters and restructures attention patterns at the proxy scale. The primary accuracy hypothesis — whether eviction-aware training improves LongBench task performance on LLaMA-2-7B and Mistral-7B-v0.1 — was not evaluated due to unavailability of gated model access. Full accuracy evaluation remains as future work, supported by the validated BudgetSweepEvaluator and SpearmanAnalyzer infrastructure described here.

