
This paper identifies the training-inference distribution mismatch in KV-cache-constrained LoRA deployment — adapters fine-tuned with full KV cache access are routinely evaluated under eviction policies that discard half or more of that cache — and proposes Eviction-Aware LoRA as a direct response. The method applies H2O eviction masks during the LoRA forward pass via a training-only hook, exposing adapter parameters to the evicted-cache attention distribution during training.

Experiments on a GPT-2 proxy model confirm two mechanistic results: (1) eviction-aware training produces adapter weights with a mean cosine similarity of 0.053 relative to sequential-baseline adapters (range: −0.578 to +0.469; all 24 LoRA layers below the 0.95 threshold) after 30 training steps; (2) attention entropy divergence is significant (paired t-test, p < 0.05) in 8 of 12 transformer layers on 5 synthetic samples, concentrated in middle layers 4–11, with no significant divergence in lower layers 0–3. Heavy-hitter concentration divergence was not significant in any layer.

The primary accuracy hypothesis — whether these mechanistic changes translate to ≥2\% per-category LongBench accuracy improvement at r=50\% on LLaMA-2-7B and Mistral-7B-v0.1 — was not tested. H-M3 is the immediate next experiment, estimated at approximately 24 GPU-hours on A100 hardware, using the validated BudgetSweepEvaluator infrastructure.

Three additional directions follow directly from this work. First, executing H-M3 will determine whether proxy-scale weight divergence predicts full-scale accuracy differences and will help distinguish between interpretations of the divergence magnitude (genuine optimization landscape divergence vs. short-training instability amplified at proxy scale). Second, the H2O+SDPA incompatibility should be addressed at the kernel level to remove the inference overhead introduced by \texttt{attn_implementation='eager'}. Third, evaluating eviction-aware adapters trained with H2O masks under SnapKV or StreamingLLM eviction at inference would test whether token-scarcity regularization generalizes across eviction policies.

