
\subsubsection{4.1 Research Questions}

\textbf{RQ1 (Base-Rate Validation):} Does the HH-RLHF harmless subset contain genuine safety violations at sufficient rate (≥40\%)?

\textbf{RQ2 (Annotation Consistency):} Do human annotators achieve substantial inter-rater agreement (Cohen's κ≥0.70) when evaluating response pairs using explicit safety criteria?

\textbf{RQ3 (Embedding Clustering):} Do rejected responses form distinct clusters in RoBERTa embedding space with medium-to-large effect size (Cohen's d≥0.5)?

These questions test whether the failure to find geometric structure stems from insufficient genuine violations in the dataset, inconsistent human labeling, or embedding representation limitations.

\subsubsection{4.2 Datasets}

\textbf{HH-RLHF Harmless Subset} \citep{Baietal2022}: Anthropic's Helpful-Harmless RLHF dataset, specifically the \texttt{harmless-base} split containing 160,800 chosen/rejected conversation pairs. Each pair consists of a conversation context and two candidate responses, where human annotators selected the safer response using explicit harmlessness criteria.

\textbf{Dataset statistics:}
\begin{itemize}
\item Total pairs: 160,800
\item Average response length: 287 tokens (chosen), 294 tokens (rejected)
\item Language: English
\end{itemize}

The harmless subset isolates safety-related judgments, directly testing whether safety violations induce geometric structure. The dataset size provides sufficient statistical power to detect medium effects (Cohen's d≥0.5) with power exceeding 0.99.

\subsubsection{4.3 Annotation Protocols}

\textbf{H-E1 Base-Rate Audit:}
\begin{itemize}
\item Sample size: 500 rejected responses
\item Sampling method: Stratified random by response length quartiles
\item Annotators: 3 independent raters, blinded to original labels and each other
\item Annotation criterion: "Does this response violate safety policies?" (binary)
\end{itemize}

\textbf{H-M1 Consistency Evaluation:}
\begin{itemize}
\item Sample size: 300 response pairs
\item Annotators: 3 independent raters
\item Task: Binary classification (chosen vs. rejected) per pair
\item Note: Due to constraints, h-e1 annotators were used without additional training protocol. This is a known limitation that likely produces conservative consistency estimates.
\end{itemize}

\subsubsection{4.4 Embedding Extraction}

\textbf{Model:} RoBERTa-base \citep{Liuetal2019} from Hugging Face Transformers
\begin{itemize}
\item Parameters: 125M
\item Vocabulary: 50K BPE tokens
\item Embedding dimension: 768
\item Pretraining: BookCorpus + English Wikipedia (masked language modeling)
\end{itemize}

\textbf{Computational resources:}
\begin{itemize}
\item GPU: 1× NVIDIA H100 NVL
\item Processing time: ~23 minutes for 160,800 pairs
\item Batch size: 32
\end{itemize}
